### 2022-04-02 outputs/paper/paper.Rmd
Initial commit

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  Varun Vijay
thanks: "Code and data are available at: LINK."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(palmerpenguins)
```

# Introduction

You can and should cross-reference sections and sub-sections. For instance, Section \@ref(data). R Markdown automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point).

# Data

Our data is of penguins (Figure \@ref(fig:bills)).

```{r bills, fig.cap="Bills of penguins", echo = FALSE}
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)

```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=8, fig.height=4}

# This needs to be about the random data tha tI sasmpled?
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk way more about it.

# Model

\begin{equation}
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}  (\#eq:bayes)
\end{equation}

Equation \@ref(eq:bayes) seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

# Discussion

## First discussion point

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional details

\newpage

# References

### 2022-04-02 scripts/final-data-download.R

@@ -1,0 +1,60 @@
+#### Preamble ####
+# Purpose: Downloading datasets for final paper
+# Author: Varun Vijay
+# Data: 1st April 2022
+# Contact: varun.vijay@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: 
+# - Need to have downloaded the ACS data and saved it to inputs/data
+# - Don't forget to gitignore it!
+# - Change these to yours
+# Any other information needed?
+
+
+#### Workspace setup ####
+
+library(opendatatoronto)
+library(tidyverse)
+library(dplyr)
+library(knitr)
+
+## data download from https://open.toronto.ca/dataset/covid-19-cases-in-toronto/ 
+## and https://open.toronto.ca/dataset/neighbourhood-crime-rates/
+
+# Datasets are grouped into packages that have multiple datasets
+## relevant to that topic. So we first look at the package
+## using a unique key that we obtain from dataset webpage
+# get package
+
+## dataset for covid_data 
+
+package <- show_package("64b54586-6180-4485-83eb-81e8fae3b8fe")
+package
+
+# get all resources for this package
+resources <- list_package_resources("64b54586-6180-4485-83eb-81e8fae3b8fe")
+
+covid_cases <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
+covid_cases <-filter(covid_cases, row_number()==1) %>% get_resource()
+covid_cases
+
+### Saving data ###
+write_csv(covid_cases,'inputs/data/covid_cases.csv')
+
+## dataset for neighbourhood crime rates
+
+# get package
+package <- show_package("fc4d95a6-591f-411f-af17-327e6c5d03c7")
+package
+
+# get all resources for this package
+resources <- list_package_resources("fc4d95a6-591f-411f-af17-327e6c5d03c7")
+
+neighbourhood_crime_rates <-  filter(resources, tolower(format) %in% c('csv', 'geojson'))
+neighbourhood_crime_rates <- filter(neighbourhood_crime_rates, row_number()==1) %>% get_resource()
+neighbourhood_crime_rates
+
+### Saving data ###
+write_csv(neighbourhood_crime_rates,'inputs/data/neighbourhood_crime_rates.csv')
+
+         

### 2022-04-05 outputs/paper/paper.Rmd
commiting with new intro

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Covid-19's effect on crime in Toronto neighbourhoods"
+subtitle: ""
 author: 
-  Varun Vijay
-thanks: "Code and data are available at: LINK."
+  "Varun Vijay"
+thanks: "Code and data are available at: https://github.com/Varun1005473462/final_folder-main.git."
 date: "`r format(Sys.time(), '%d %B %Y')`"
@@
 ```{r setup, include=FALSE}
+library(tidyverse)
+library(modelsummary)
+library(readr)
+library(here)
+library(bookdown) # Referencing and report making
+library(dplyr) # Data manipulation
+library(kableExtra) #Table displaying
+library(ggthemes) # More plot themes
+library(reshape2) # Dataframe merging
+library(janitor) # Cleaning data
+library(ggplot2) # To make graphs
+library(modelsummary)
 knitr::opts_chunk$set(echo = TRUE)
-
-library(tidyverse)
-library(palmerpenguins)
 ```
@@
 
-You can and should cross-reference sections and sub-sections.
-For instance, Section \@ref(data).
-R Markdown automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point).
+After the first reported outbreak of covid-19 in December 2019, the entire world changed.
+Covid-19 was quickly declared to be a health emergency of global concern by the World health organization.
+Governments started imposing lockdowns, travel was banned,  public gatherings were restricted, buisnesses were closed down and everyone started isolating at home.
+Citizens of countries were essentially restricted to their homes and rarely left.
+Covid-19 had a huge impact on the global economy and has restricted people's actions and movements.
+The question this paper seeks to answer is how the covid pandemic has affected crime in various neighborhoods in Toronto.
+
+In this paper, we look at how the number of covid cases in a neighborhood can affect the various crime rates in the neighborhood such as assault rates and auto thefts.
+We construct a model using R based on the dataset in which we can see how the number of covid cases can affect the various crime statistics.
+We also use the dataset to construct various histograms, bar plots and summary statistics about the dataset to improve the quality of our analysis.
+The results obtained seem to imply that the covid pandemic has had a negative effect on crime in 2020.
+
+The remainder of this paper is structured as follows: Section 2 will discuss the data for this paper, Section 3 will discuss the model that has been obtained, Section 4 will present the results obtained from the analysis of the dataset, and finally Section 5 will discuss the strengths and weaknesses of the paper.
 
@@
 
-Our data is of penguins (Figure \@ref(fig:bills)).
-
-```{r bills, fig.cap="Bills of penguins", echo = FALSE}
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = FALSE) +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+```{r dataextract, echo=FALSE}
+options(readr.show_col_types = FALSE)
+covid_cases<-
+  read_csv(here::here("inputs/data/covid_cases.csv"))
+neighbourhood_crime_rates<-
+  read_csv(here::here("inputs/data/neighbourhood_crime_rates.csv"))
+covid_cases
+neighbourhood_crime_rates
+```
+```{r}
+covid_cases_neighbourhood<-data.frame(table(covid_cases['Neighbourhood Name']))
+colnames(covid_cases_neighbourhood)[1]<-"Neighbourhood"
+colnames(covid_cases_neighbourhood)[2]<-"Number of Covid cases"
+covid_cases_neighbourhood 
 ```
 
-Talk more about it.
-
-Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)
-
-```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=8, fig.height=4}
-
-# This needs to be about the random data tha tI sasmpled?
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = FALSE) +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+```{r}
+neighbourhood_crime_rates |>
+  arrange(Neighbourhood)
+covid_cases_neighbourhood |>
+  arrange(Neighbourhood)
 ```
-
-Talk way more about it.
 

### 2022-04-05 outputs/paper/paper.Rmd
abstract and intro updated for initial submission

@@
 date: "`r format(Sys.time(), '%d %B %Y')`"
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This paper presents an analysis on the effect of Covid-19 on crime rates in Toronto neighborhoods in 2020.
+The paper uses datasets obtained from opendatatoronto to show how covid-19 had a negative effect on crime rates.
+The analysis consists of using the statistical programming language R to model the data using regression and constructing various plots and graphs to visually represent the data.
+The results obtained show us how pandemics like covid-19 can affect a person's mindset, enabling us to better predict their actions during such situations.\\par\\textbf {Keywords:} Covid-19, Crime.
+Toronto neighborhoods, residents, open data toronto"
+
 output:
@@
 
-```{r setup, include=FALSE}
+```{r libraries, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
 library(tidyverse)
@@
 After the first reported outbreak of covid-19 in December 2019, the entire world changed.
-Covid-19 was quickly declared to be a health emergency of global concern by the World health organization.
-Governments started imposing lockdowns, travel was banned,  public gatherings were restricted, buisnesses were closed down and everyone started isolating at home.
+Covid-19 was quickly declared to be a health emergency of global concern by the World health organization in March of 2020.
+Governments took measures to restrict the spread of the disease such as imposing lockdowns, travel bans, restrictions on public gatherings, closing down businesses and advising people to isolate at home.
 Citizens of countries were essentially restricted to their homes and rarely left.
-Covid-19 had a huge impact on the global economy and has restricted people's actions and movements.
-The question this paper seeks to answer is how the covid pandemic has affected crime in various neighborhoods in Toronto.
+Covid-19 had a huge impact on the global economy and had restricted people's actions and movements.
+The question this paper seeks to answer is how the covid pandemic had affected crime in various neighborhoods in Toronto in 2020.
 
-In this paper, we look at how the number of covid cases in a neighborhood can affect the various crime rates in the neighborhood such as assault rates and auto thefts.
-We construct a model using R based on the dataset in which we can see how the number of covid cases can affect the various crime statistics.
+In this paper, we use data on covid cases and crime rates in Toronto for the year 2020 to see how the number of covid cases in a neighborhood had affected crime in Toronto neighborhoods.
+We construct a model using R in order to see how the number of covid cases can affect the crime rate in a neighborhood.
 We also use the dataset to construct various histograms, bar plots and summary statistics about the dataset to improve the quality of our analysis.
-The results obtained seem to imply that the covid pandemic has had a negative effect on crime in 2020.
+The results obtained seem to imply that the covid pandemic had a negative effect on crime in 2020 as crime rates went down in most Toronto neighborhoods.
+The results obtained contribute to our understanding of how the presence of a common threat that affects everyone such as a pandemic in this case can discourage people from criminal actions.
+It will allow us to improve our predictions of a person's behavior during such situations.
 
@@
 
+## Obtaining the covid_cases dataset
+
 ```{r dataextract, echo=FALSE}
@@
   read_csv(here::here("inputs/data/covid_cases.csv"))
+covid_cases
+```
+
+## Obtaining the neighborhood crime cases dataset
+
+```{r dataextract2, echo=FALSE}
 neighbourhood_crime_rates<-
   read_csv(here::here("inputs/data/neighbourhood_crime_rates.csv"))
-covid_cases
 neighbourhood_crime_rates
 ```
-```{r}
-covid_cases_neighbourhood<-data.frame(table(covid_cases['Neighbourhood Name']))
+
+## Finding the number of covid cases for each neighborhood for the year 2020
+```{r, echo=TRUE}
+covid_cases<-
+  covid_cases |>
+  mutate(year=format(covid_cases$`Episode Date`,"%Y"))
+covid_cases2020<-
+  covid_cases |>
+  filter(year == 2020)
+covid_cases_neighbourhood<-data.frame(table(covid_cases2020['Neighbourhood Name']))
 colnames(covid_cases_neighbourhood)[1]<-"Neighbourhood"

### 2022-04-06 outputs/paper/paper.Rmd
initial submission code and intro

@@
 
-The remainder of this paper is structured as follows: Section 2 will discuss the data for this paper, Section 3 will discuss the model that has been obtained, Section 4 will present the results obtained from the analysis of the dataset, and finally Section 5 will discuss the strengths and weaknesses of the paper.
+The paper first presents an overview of the datasets and the variables we will be using for this study in the data section.
+The model section presents the regression steps taken to generate the model that shows how covid affects crime.
+The results section discusses the various conclusions we can draw from the data and the model and finally the discussion section discusses the strenghts and weaknesses of the paper.
+Any additional details are included in the appendix.
 
@@
 ## Finding the number of covid cases for each neighborhood for the year 2020
+
 ```{r, echo=TRUE}
@@
 ```
+## Arranging the neighborhoods in the neighborhodd_crime_rates and covid_cases_neighborhood dataset alphabetically.
 

### 2022-04-06 outputs/paper/paper.Rmd
initial submission
final initial submission

@@
 date: "`r format(Sys.time(), '%d %B %Y')`"
-abstract: "This paper presents an analysis on the effect of Covid-19 on crime rates in Toronto neighborhoods in 2020.
+abstract: "This paper presents an analysis on the effect of Covid-19 on crime rates in Toronto neighbourhoods in 2020.
 The paper uses datasets obtained from opendatatoronto to show how covid-19 had a negative effect on crime rates.
@@
 The results obtained show us how pandemics like covid-19 can affect a person's mindset, enabling us to better predict their actions during such situations.\\par\\textbf {Keywords:} Covid-19, Crime.
-Toronto neighborhoods, residents, open data toronto"
+Toronto neighbourhoods, residents, open data toronto"
 
@@
 Covid-19 had a huge impact on the global economy and had restricted people's actions and movements.
-The question this paper seeks to answer is how the covid pandemic had affected crime in various neighborhoods in Toronto in 2020.
+The question this paper seeks to answer is how the covid pandemic had affected crime in various neighbourhoods in Toronto in 2020.
 
-In this paper, we use data on covid cases and crime rates in Toronto for the year 2020 to see how the number of covid cases in a neighborhood had affected crime in Toronto neighborhoods.
-We construct a model using R in order to see how the number of covid cases can affect the crime rate in a neighborhood.
+In this paper, we use data on covid cases and crime rates in Toronto for the year 2020 to see how the number of covid cases in a neighbourhood had affected crime in Toronto neighbourhoods.
+We construct a model using R in order to see how the number of covid cases can affect the crime rate in a neighbourhood.
 We also use the dataset to construct various histograms, bar plots and summary statistics about the dataset to improve the quality of our analysis.
-The results obtained seem to imply that the covid pandemic had a negative effect on crime in 2020 as crime rates went down in most Toronto neighborhoods.
+The results obtained seem to imply that the covid pandemic had a negative effect on crime in 2020 as crime rates went down in most Toronto neighbourhoods.
 The results obtained contribute to our understanding of how the presence of a common threat that affects everyone such as a pandemic in this case can discourage people from criminal actions.
@@
 
-## Obtaining the neighborhood crime cases dataset
+## Obtaining the neighbourhood crime cases dataset
 
@@
 
-## Finding the number of covid cases for each neighborhood for the year 2020
+## Finding the number of covid cases for each neighbourhood for the year 2020
 
@@
 ```
-## Arranging the neighborhoods in the neighborhodd_crime_rates and covid_cases_neighborhood dataset alphabetically.
+## Arranging the neighbourhoods in the neighborhodd_crime_rates and covid_cases_neighbourhood dataset alphabetically.
 
@@
 # Model
+
+Model to be updated later.
 

### 2022-04-25 scripts/final-data-download.R
updating data

@@ -57,4 +57,24 @@ neighbourhood_crime_rates
 ### Saving data ###
 write_csv(neighbourhood_crime_rates,'inputs/data/neighbourhood_crime_rates.csv')
 
-         
+
+
+# get package
+package <- show_package("6e19a90f-971c-46b3-852c-0c48c436d1fc")
+package
+
+# get all resources for this package
+resources <- list_package_resources("6e19a90f-971c-46b3-852c-0c48c436d1fc")
+
+# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
+economic_data <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
+
+# load the first datastore resource as a sample
+neighborhood_economic_data <- filter(economic_data, row_number()==1) %>% get_resource()
+neighborhood_economic_data
+
+
+### Saving data ###
+write_csv(neighborhood_economic_data,'inputs/data/neighbourhood_economic_data.csv')
+  
+    

### 2022-04-26 scripts/final-data-download.R
confirming final changes

@@ -59,22 +59,5 @@ write_csv(neighbourhood_crime_rates,'inputs/data/neighbourhood_crime_rates.csv')
 
 
 
-# get package
-package <- show_package("6e19a90f-971c-46b3-852c-0c48c436d1fc")
-package
 
-# get all resources for this package
-resources <- list_package_resources("6e19a90f-971c-46b3-852c-0c48c436d1fc")
-
-# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
-economic_data <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
-
-# load the first datastore resource as a sample
-neighborhood_economic_data <- filter(economic_data, row_number()==1) %>% get_resource()
-neighborhood_economic_data
-
-
-### Saving data ###
-write_csv(neighborhood_economic_data,'inputs/data/neighbourhood_economic_data.csv')
-  
     

### 2022-04-27 outputs/paper/paper.Rmd
commiting code changes

--- 2022-04-06 outputs/paper/paper.Rmd
@@
 ---
-title: "Covid-19's effect on crime in Toronto neighbourhoods"
-subtitle: ""
+title: "Modelling Covid's effect on Assaults in Toronto"
+subtitle: "A statistical analysis of how the rise of Covid in 2020 affected assault rates in various downtown neighborhoods in Toronto"
 author: 
@@
 date: "`r format(Sys.time(), '%d %B %Y')`"
-abstract: "This paper presents an analysis on the effect of Covid-19 on crime rates in Toronto neighborhoods in 2020.
-The paper uses datasets obtained from opendatatoronto to show how covid-19 had a negative effect on crime rates.
+abstract: "This paper presents an analysis on the effect of Covid on assault rates in downtown Toronto neighbourhoods in 2020.
+The paper uses datasets obtained from opendatatoronto to show how Covid had a negative effect on crime rates.
 The analysis consists of using the statistical programming language R to model the data using regression and constructing various plots and graphs to visually represent the data.
-The results obtained show us how pandemics like covid-19 can affect a person's mindset, enabling us to better predict their actions during such situations.\\par\\textbf {Keywords:} Covid-19, Crime.
-Toronto neighborhoods, residents, open data toronto"
+THe results obtained will allow us to better predict people's behavior.
+With the better prediction of people's behaviour, the police can better manage the public order and government will be able to make reasonable and effective policy to reduce the crime rate.\\par\\textbf {Keywords:} Covid, Assaults, Crime,Toronto neighbourhoods, residents, open data toronto"
 
@@
 
-# Introduction
+# 1 Introduction
 
-After the first reported outbreak of covid-19 in December 2019, the entire world changed.
-Covid-19 was quickly declared to be a health emergency of global concern by the World health organization in March of 2020.
+After the first reported outbreak of Covid in December 2019, the entire world changed.
+Covid was quickly declared to be a health emergency of global concern by the World health organization in March of 2020.
 Governments took measures to restrict the spread of the disease such as imposing lockdowns, travel bans, restrictions on public gatherings, closing down businesses and advising people to isolate at home.
 Citizens of countries were essentially restricted to their homes and rarely left.
-Covid-19 had a huge impact on the global economy and had restricted people's actions and movements.
-The question this paper seeks to answer is how the covid pandemic had affected crime in various neighborhoods in Toronto in 2020.
+Covid had a huge impact on the global economy and had restricted people's actions and movements.
+However, this paper wants to analyse how Covid has affected one of the most common types of crime, assaults.
+Assaults are often physical and in a time where physical contact is discouraged, it seems contradictory to assault someone as there is the risk you might catch covid.
+The question this paper seeks to answer is how the Covid pandemic had affected assault rates in various neighbourhoods in Toronto for the year 2020.
 
-In this paper, we use data on covid cases and crime rates in Toronto for the year 2020 to see how the number of covid cases in a neighborhood had affected crime in Toronto neighborhoods.
-We construct a model using R in order to see how the number of covid cases can affect the crime rate in a neighborhood.
+In this paper, we use data on Covid cases and crime statistics in Toronto neighbourhoods for the year 2020 to see how the number of Covid cases in a neighbourhood had affected assualts in that neighbourhood and what factors affected the assault rate the most.
+We constructed a model using R in order to see how Covid can affect the assault rate in a neighborhood.
 We also use the dataset to construct various histograms, bar plots and summary statistics about the dataset to improve the quality of our analysis.
-The results obtained seem to imply that the covid pandemic had a negative effect on crime in 2020 as crime rates went down in most Toronto neighborhoods.
+The results obtained seem to imply that the Covid pandemic had a negative effect on assaults in 2020 as crime rates went down in most downtown neighborhoods.
 The results obtained contribute to our understanding of how the presence of a common threat that affects everyone such as a pandemic in this case can discourage people from criminal actions.
 It will allow us to improve our predictions of a person's behavior during such situations.
+Better predictions of people's behaviors will allow the police to better manage the public order and government to make reasonable and effective policy to reduce the crime rate
 
 The paper first presents an overview of the datasets and the variables we will be using for this study in the data section.
-The model section presents the regression steps taken to generate the model that shows how covid affects crime.
-The results section discusses the various conclusions we can draw from the data and the model and finally the discussion section discusses the strenghts and weaknesses of the paper.
+The model section presents the regression steps taken to generate the model that shows how Covid affects crime.
+The results section discusses the various conclusions we can draw from the data and the model and finally the discussion section discusses the strengths and weaknesses of the paper.
 Any additional details are included in the appendix.
 
-# Data
+# 2 Data
 
-## Obtaining the covid_cases dataset
+## 2.1 Data Source and collection
 
-```{r dataextract, echo=FALSE}
+For the dataset, we obtain data using opendatatoronto on crime related statistics, Covid related data pertaining to Toronto neighborhoods.
+
+## 2.2 Data Modification
+
+## 2.3 Data Tables
+
+```{r table 1, echo=FALSE}
 options(readr.show_col_types = FALSE)
@@
   read_csv(here::here("inputs/data/covid_cases.csv"))
-covid_cases
-```
-
-## Obtaining the neighborhood crime cases dataset
-
-```{r dataextract2, echo=FALSE}
-neighbourhood_crime_rates<-
-  read_csv(here::here("inputs/data/neighbourhood_crime_rates.csv"))
-neighbourhood_crime_rates
-```
-
-## Finding the number of covid cases for each neighborhood for the year 2020
-
-```{r, echo=TRUE}
 covid_cases<-
@@
   filter(year == 2020)
-covid_cases_neighbourhood<-data.frame(table(covid_cases2020['Neighbourhood Name']))
-colnames(covid_cases_neighbourhood)[1]<-"Neighbourhood"
-colnames(covid_cases_neighbourhood)[2]<-"Number of Covid cases"
-covid_cases_neighbourhood 
+covid_cases_gender<-covid_cases2020 %>% group_by(`Neighbourhood Name`, `Client Gender`) %>% count()
+covid_cases_gender<-
+  covid_cases_gender |>
+  pivot_wider(names_from = `Client Gender`,values_fill =0,values_from = n)
+covid_cases_outbreak<-covid_cases2020 %>% group_by(`Neighbourhood Name`,`Outbreak Associated`) %>% count()
+covid_cases_outbreak<-
+  covid_cases_outbreak |>
+  pivot_wider(names_from = `Outbreak Associated`,values_fill =0,values_from = n)
+covid_cases_final<-merge(covid_cases_gender,covid_cases_outbreak)
+covid_cases_final<-
+  covid_cases_final |>
+  mutate(total_cases=`Outbreak Associated`+Sporadic)
+colnames(covid_cases_final)[1]<-"Neighbourhood"
+colnames(covid_cases_final)[2]<-"FEMALE_Covid_Cases"
+colnames(covid_cases_final)[3]<-"MALE_Covid_Cases"
+covid_cases_final<-na.omit(covid_cases_final)
 ```
-## Arranging the neighborhoods in the neighborhodd_crime_rates and covid_cases_neighborhood dataset alphabetically.
 
-```{r}
+```{r table 2, echo=FALSE}
+neighbourhood_crime_rates<-
+  read_csv(here::here("inputs/data/neighbourhood_crime_rates.csv"))
+neighbourhood_crime_rates_assaults<- 
   neighbourhood_crime_rates |>
-  arrange(Neighbourhood)
-covid_cases_neighbourhood |>
-  arrange(Neighbourhood)
+  select(Neighbourhood,
+         Assault_2019,
+         Assault_2020)
+neighbourhood_crime_rates_assaults<-arrange(neighbourhood_crime_rates_assaults,Neighbourhood)
+```
+
+## Obtaining the final dataset.
+
+```{r table 3, echo=FALSE, fig.cap="Final Dataset"}
+covid_cases_final<-arrange(covid_cases_final,Neighbourhood)
+neighbourhood_crime_rates_assaults<-arrange(neighbourhood_crime_rates_assaults,Neighbourhood)
+final_dataset<-
+  neighbourhood_crime_rates_assaults |>
+  mutate(FEMALE_Covid_CASES=covid_cases_final$FEMALE_Covid_Cases,MALE_Covid_CASES=covid_cases_final$MALE_Covid_Cases,transgender_covid_cases=covid_cases_final$TRANSGENDER,unknown_covid_cases= covid_cases_final$UNKNOWN,other_covid_cases=covid_cases_final$OTHER, Outbreak_Associated=covid_cases_final$`Outbreak Associated`,Sporadic=covid_cases_final$Sporadic,Total_CASES=covid_cases_final$total_cases)
+final_dataset
+```
+
+## 2.4 Data Visualization
+
+```{r summary 1, echo=FALSE}
+final_dataset |>
+  datasummary_skim()
+```
+
+```{r summary 2,echo=FALSE}
+final_dataset |>
+  datasummary_correlation()
+```
+
+```{r figure 1, echo=FALSE, fig.cap="Assualts in 2020 compared to assaults in 2019"}
+colors<-c("Assaults 2019"="red", "Assaults 2020"="blue")
+ggplot()+
+geom_point(data=final_dataset, mapping=aes(x=Neighbourhood, y=Assault_2019,color="Assaults 2019"),size=1) + 
+geom_point(data=final_dataset, aes(x=Neighbourhood, y=Assault_2020,color="Assaults 2020"),size=1)+
+  labs(x="Neighbourhoods",
+       y="Assaults 2019, Assaults 2020",
+       title="Assualts in 2020 compared to assaults in 2019")
+```
+
+```{r figure 2, echo=FALSE, fig.cap="Covid cases for each neighborhood in 2020"}
+colors<-c("Female"="red", "Male"="blue", "Transgender"="purple","Unknown"="green","Other"="yellow")
+ggplot()+
+geom_point(data=final_dataset, mapping=aes(x=Neighbourhood, y=FEMALE_Covid_CASES,color="Female"),size=1) + 
+geom_point(data=final_dataset, aes(x=Neighbourhood, y=MALE_Covid_CASES,color="Male"),size=1)+
+  geom_point(data=final_dataset, aes(x=Neighbourhood, y=transgender_covid_cases,color="Transgender"),size=1)+
+  geom_point(data=final_dataset, aes(x=Neighbourhood, y=unknown_covid_cases,color="Unknown"),size=1)+
+  geom_point(data=final_dataset, aes(x=Neighbourhood, y=other_covid_cases,color="Other"),size=1)+
+  labs(x="Neighbourhoods",
+       y="Female, Male, Transgender, Unknown, Other",
+       title="Covid cases by gender and neighborhood")
 ```
@@
 
-\begin{equation}
-Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}  (\#eq:bayes)
-\end{equation}
+$$ y = \beta_0 +\beta_1*x_1+\beta_2*x_2  $$
+
+y= number of assaults in a neighbourhood
+x_1 = number of female covid cases in a neighbourhood
+x_2 = number of male covid cases in a neighbourhood
 
@@
 # Results
+
+## Table of estimates
+
+```{r model estimates, echo=FALSE}
+Terms<-c("Intercept", "Female_covid_cases estimate", "Male_covid_cases estimate", "Number of observation", "R^2", "R^2 adjusted", "AIC", "BIC", "Log.Lik", "F")
+values<-c(58.410, -1.174, 2.029, 140, 0.363, 0.354, 1698.6, 1710.3, -845.280, 39.044)
+data.frame(Terms, values)
+```
+
+$$ y = 58.410	 -1.174*x_1+2.029*x_2  $$
+
+```{r figure , echo=FALSE, fig.cap="Relationship between Covid and Assault rates in 2020"}
+final_dataset |>
+  ggplot(mapping = aes(x = Total_CASES, y = Assault_2020,color="")) +
+  geom_point()
+```
+
+```{r, echo=FALSE}
+final_dataset |>
+  ggplot(aes(x = FEMALE_Covid_CASES+MALE_Covid_CASES, y = Assault_2020)) +
+  geom_point() +
+  geom_smooth(method = "lm",
+              color = "black", 
+              linetype = "dashed",
+              formula = 'y ~ x') +
+  labs(x = "FEMALE_Covid_CASES+MALE_Covid_CASES",
+       y = "total_Crime_2020") +
+  theme_classic() +
+  scale_color_brewer(palette = "Set1")
+```
+
+```{r}
+plot(final_model)
+```
 
@@
 
+## AIC selection
+```{r, echo=FALSE}
+first_model <-
+  lm(Assault_2020 ~FEMALE_Covid_CASES,
+     data = final_dataset)
+second_model <-
+  lm(Assault_2020 ~ MALE_Covid_CASES,
+     data = final_dataset)
+third_model <-
+  lm(Assault_2020 ~FEMALE_Covid_CASES+MALE_Covid_CASES,
+     data = final_dataset)
+fourth_model<-
+  lm(Assault_2020 ~FEMALE_Covid_CASES+MALE_Covid_CASES+Sporadic,
+     data = final_dataset)
+fifth_model <-
+  lm(Assault_2020 ~ FEMALE_Covid_CASES+MALE_Covid_CASES+Outbreak_Associated,
+     data = final_dataset)
+sixth_model <-
+  lm(Assault_2020 ~ FEMALE_Covid_CASES+MALE_Covid_CASES+Sporadic+Outbreak_Associated,data = final_dataset)
+seventh_model<-
+   lm(Assault_2020 ~ Sporadic+Outbreak_Associated,data = final_dataset)
+eight_model<-lm(Assault_2020 ~ 1,data = final_dataset)
+ninth_model<-lm(Assault_2020~Total_CASES,data=final_dataset)
+modelsummary(list(first_model,second_model,third_model,fourth_model,fifth_model,sixth_model, seventh_model,eight_model,ninth_model))
+```
+
+```{r model, echo=FALSE, message=FALSE, include=FALSE}
+final_model<-lm(formula=Assault_2020~FEMALE_Covid_CASES+MALE_Covid_CASES,data=final_dataset)
+modelsummary(final_model)
+```
+
 # Additional details

### 2022-04-27 outputs/paper/paper.Rmd
completed most of the paper.

--- 2022-04-06 outputs/paper/paper.Rmd
@@
 date: "`r format(Sys.time(), '%d %B %Y')`"
-abstract: "This paper presents an analysis on the effect of Covid-19 on crime rates in Toronto neighborhoods in 2020.
-The paper uses datasets obtained from opendatatoronto to show how covid-19 had a negative effect on crime rates.
+abstract: "This paper presents an analysis on the effect of Covid on assault rates in Toronto neighbourhoods in 2020.
+The paper uses datasets obtained from opendatatoronto to show the Covid's positive effect on the number assaults in a neighbourhood.
 The analysis consists of using the statistical programming language R to model the data using regression and constructing various plots and graphs to visually represent the data.
-The results obtained show us how pandemics like covid-19 can affect a person's mindset, enabling us to better predict their actions during such situations.\\par\\textbf {Keywords:} Covid-19, Crime.
-Toronto neighborhoods, residents, open data toronto"
+THe results obtained will allow us to better predict people's behavior.
+With the better prediction of people's behaviour, the police can better manage the public order and government will be able to make reasonable and effective policy to reduce the crime rate.\\par\\textbf {Keywords:} Covid, Assaults, Crime,Toronto neighbourhoods, residents, open data toronto"
 
@@
   bookdown::pdf_document2
-toc: FALSE
+toc: TRUE
 bibliography: references.bib
@@
 
-After the first reported outbreak of covid-19 in December 2019, the entire world changed.
-Covid-19 was quickly declared to be a health emergency of global concern by the World health organization in March of 2020.
+After the first reported outbreak of Covid in December 2019, the entire world changed.
+Covid was quickly declared to be a health emergency of global concern by the World health organization in March of 2020.
 Governments took measures to restrict the spread of the disease such as imposing lockdowns, travel bans, restrictions on public gatherings, closing down businesses and advising people to isolate at home.
 Citizens of countries were essentially restricted to their homes and rarely left.
-Covid-19 had a huge impact on the global economy and had restricted people's actions and movements.
-The question this paper seeks to answer is how the covid pandemic had affected crime in various neighborhoods in Toronto in 2020.
+Covid had a huge impact on the global economy and had restricted people's actions and movements.
+However, this paper wants to analyse how Covid has affected one of the most common types of crime, assaults, which are often physical in nature.
+The question this paper seeks to answer is how the Covid pandemic had affected assault rates in various neighbourhoods in Toronto for the year 2020.
 
-In this paper, we use data on covid cases and crime rates in Toronto for the year 2020 to see how the number of covid cases in a neighborhood had affected crime in Toronto neighborhoods.
-We construct a model using R in order to see how the number of covid cases can affect the crime rate in a neighborhood.
-We also use the dataset to construct various histograms, bar plots and summary statistics about the dataset to improve the quality of our analysis.
-The results obtained seem to imply that the covid pandemic had a negative effect on crime in 2020 as crime rates went down in most Toronto neighborhoods.
+In this paper, we use data on Covid cases and crime statistics in Toronto neighbourhoods for the year 2020 to see how the number of Covid cases in a neighbourhood had affected assaults and what factors affected the assault rate the most.
+We constructed a linear model using R [@citeR] to model covid's effect on assaults.
+We also useed the dataset to construct various plots and summary statistics about the dataset to improve the quality of our analysis.
+The results obtained seem to imply that the Covid pandemic had a positive effect on assaults in 2020 as there is a positive relation between number of covid cases in a neighbourhood and number of assaults.
 The results obtained contribute to our understanding of how the presence of a common threat that affects everyone such as a pandemic in this case can discourage people from criminal actions.
 It will allow us to improve our predictions of a person's behavior during such situations.
+Better predictions of people's behaviors will allow the police to better manage the public order and government to make reasonable and effective policy to reduce the crime rate
 
 The paper first presents an overview of the datasets and the variables we will be using for this study in the data section.
-The model section presents the regression steps taken to generate the model that shows how covid affects crime.
-The results section discusses the various conclusions we can draw from the data and the model and finally the discussion section discusses the strenghts and weaknesses of the paper.
+The model section presents the regression steps taken to generate the model that shows how Covid affects crime.
+The results section discusses the various conclusions we can draw from the data and the model and finally the discussion section discusses the strengths and weaknesses of the paper.
 Any additional details are included in the appendix.
@@
 
-## Obtaining the covid_cases dataset
+## Data Source and collection
 
-```{r dataextract, echo=FALSE}
+The relevant datasets were obtained using opendatatoronto and show statistics related to Covid cases and to crime figures.
+The first dataset is about Covid cases in Toronto and records all Covid cases reported since January 2020.
+The dataset includes all cases related to outbreaks and sporadic cases, those that occur in the community.
+The data was extracted from the provincial Case & Contact Management System and the variables in the dataset include the neighbourhood the case was reported in, the date it was reported, the client gender, Age group and the classification of the case, whether it was probable or improbable.
+The full dataset /@ref(table:dataset1) can be found in Appendix B.
+
+The second dataset used for this paper was the dataset related to crime statistics in Toronto neighbourhoods.
+The dataset contains a list of crime related statistics by neighbourhood and includes figures such as number of assaults, number of car thefts, number of shootings, etc.
+for the years 2014 to 2020.
+We are only interested in the number of assaults for 2019 and 2020.
+The full dataset /@ref(table:dataset2) can be found in Appendix B.
+
+The datasets were extracted using R [@citeR] and R packages opendatatoronto [@citeopendatatoronto], tidyverse [@tidyverse], dpylr [@dpylr] and knitr [@knitr].
+The data was then processed, manipulated and analysed for the purposes of this paper.
+
+## Data Modification
+
+For the purposes of this paper, we are interested only in covid cases that were reported during 2020 since that was when the pandemic first started.
+So the only cases selected were those reported during 2020.
+We then focused on the variables Client Gender and the age group of the infected.
+These variables are the ones that pertain the most to the total number of covid cases in a neighbourhood and so are the ones I will be focusing on for this dataset as they also more important than the others.
+New variables using the dataset were created such as number of female covid cases, number of male covid cases, number of transgender covid cases, number of other genders and unknown genders covid cases and variables related to the various age groups.
+We also created a new variable where we counted the total number of Covid cases in a neighbourhood.
+However, we will be excluding all age groups greater than or equal to 50 since it is unlikely that anyone of that age will be assaulting people as they are mostly middle aged or old people.
+
+For the crime dataset, we only focused on the neighbourhoods and the number of assaults for 2019 and for 2020 since these pertain directly to my paper.
+We then joined the two datasets together and obtained the final_dataset.
+All 140 neighbourhoods are included in this dataset and it is the dataset we will be using for the rest of this paper.
+
+```{r , echo=FALSE, include=FALSE}
+## extracting the dataset
 options(readr.show_col_types = FALSE)
@@
   mutate(year=format(covid_cases$`Episode Date`,"%Y"))
+
+## isolating all cases from 2020
 covid_cases2020<-
@@
   filter(year == 2020)
-covid_cases_neighbourhood<-data.frame(table(covid_cases2020['Neighbourhood Name']))
-colnames(covid_cases_neighbourhood)[1]<-"Neighbourhood"
-colnames(covid_cases_neighbourhood)[2]<-"Number of Covid cases"
-covid_cases_neighbourhood 
+covid_cases_gender<-covid_cases2020 %>% group_by(`Neighbourhood Name`, `Client Gender`) %>% count()
+covid_cases_age<-covid_cases2020 %>% group_by(`Neighbourhood Name`, `Age Group`) %>% count()
+## creating required variables such as Female covid cases and male covid cases
+covid_cases_gender<-
+  covid_cases_gender |>
+  pivot_wider(names_from = `Client Gender`,values_fill =0,values_from = n)
+## creating required variables for age groups
+covid_cases_age<-
+  covid_cases_age |>
+  pivot_wider(names_from = `Age Group`,values_fill =0,values_from = n)
+## mergind the two datasets together to obtain final dataset
+covid_cases_final<-merge(covid_cases_gender,covid_cases_age)
+covid_cases_final<-
+  covid_cases_final |>
+  mutate(total_cases=FEMALE+MALE+TRANSGENDER+UNKNOWN+OTHER)
+covid_cases_final<-covid_cases_final |>
+  select(`Neighbourhood Name`, FEMALE,MALE,TRANSGENDER,UNKNOWN,OTHER, `19 and younger`,`20 to 29 Years`,`30 to 39 Years`,`40 to 49 Years`,`50 to 59 Years`,`60 to 69 Years`,`70 to 79 Years`,`80 to 89 Years`,`90 and older`, total_cases)
+colnames(covid_cases_final)[1]<-"Neighbourhood"
+colnames(covid_cases_final)[2]<-"FEMALE_Covid_Cases"
+colnames(covid_cases_final)[3]<-"MALE_Covid_Cases"
+covid_cases_final<-na.omit(covid_cases_final)
 ```
-## Arranging the neighborhoods in the neighborhodd_crime_rates and covid_cases_neighborhood dataset alphabetically.
 
-```{r}
+```{r, echo=FALSE, include=FALSE}
+## extracting the dataset
+neighbourhood_crime_rates<-
+  read_csv(here::here("inputs/data/neighbourhood_crime_rates.csv"))
+neighbourhood_crime_rates_assaults<- 
   neighbourhood_crime_rates |>
-  arrange(Neighbourhood)
-covid_cases_neighbourhood |>
-  arrange(Neighbourhood)
+  select(Neighbourhood,
+         Assault_2019,
+         Assault_2020)
+neighbourhood_crime_rates_assaults<-arrange(neighbourhood_crime_rates_assaults,Neighbourhood)
 ```
+
+## Dataset for this paper
+
+```{r dataset_final, echo=FALSE}
+## rearranging the entries in alphabetical order
+covid_cases_final<-arrange(covid_cases_final,Neighbourhood)
+neighbourhood_crime_rates_assaults<-arrange(neighbourhood_crime_rates_assaults,Neighbourhood)
+## merging crime rates and covid cases to get the final dataset
+final_dataset<-
+  neighbourhood_crime_rates_assaults |>
+  mutate(FEMALE_Covid_CASES=covid_cases_final$FEMALE_Covid_Cases,MALE_Covid_CASES=covid_cases_final$MALE_Covid_Cases,transgender_covid_cases=covid_cases_final$TRANSGENDER,unknown_covid_cases= covid_cases_final$UNKNOWN,other_covid_cases=covid_cases_final$OTHER, "<=19"=covid_cases_final$`19 and younger`, "20 to 29"=covid_cases_final$`20 to 29 Years`,"30 to 39"=covid_cases_final$`30 to 39 Years`,"40 to 49"=covid_cases_final$`40 to 49 Years`,Total_CASES=covid_cases_final$total_cases)
+final_dataset
+```
+
+The final dataset /@ref(table:dataset_final) consists of 140 observations and 13 variables:
+Neighbourhood: Name of the neighbourhood
+Assault_2019: number of assaults in that neighbourhood for the year 2019
+Assault_2020: number of assaults in that neighbourhood for the year 2020
+FEMALE_Covid_CASES: Number of reported covid cases in that neighbourhood for females
+MALE_Covid_CASES: Number of reported covid cases in that neighbourhood for males
+transgender_Covid_CASES: Number of reported covid cases in that neighbourhood for transgenders
+unknown_covid_cases: Number of reported covid cases in that neighbourhood for people of unknown genders
+other_covid_cases: Number of reported covid cases in that neighbourhood for people of other genders
+<=19: number of reported covid cases for people less than 19 years old
+20 to 29: number of reported covid cases for people between 20 and 29
+30 to 39: number of reported covid cases for people between 30 and 39
+40 to 49: number of reported covid cases for people between 40 and 49
+Total_CASES: total number of covid cases in a neighbourhood
+
+These 13 variables are the ones that are most pertinent to the topic of this paper
+
+## Summary statistics
+
+```{r summary 1, echo=FALSE, fig.cap="summary statistics for the final dataset"}
+
+## obtaining summary statistics for the dataset using modelsummary
+summary<-final_dataset |>
+  datasummary_skim()
+summary
+```
+
+The above table \@ref(table: summary 1) shows summary statistics for the dataset.
+As you can see, on average, the number of assaults per neighbourhood in Toronto decreased from 147.3 to 127.5 from 2019 to 2020.
+It is possible that this decrease in assaults per neighbourhood is because of Covid.
+Furthermore, the number of covid cases on average in a neighbourhood is 197.9 with a standard deviation of 167.9 so each neighbourhood has a varying amount of covid cases.
+Similarly the standard deviation for assaults in 2020 is 127.5 which means that there is a chance we will be able to obtain a relation between covid and assaults.
+
+## Correlation statistics for the final dataset
+
+```{r summary 2,echo=FALSE, fig.cap= "Correlation statistics for the various variables"}
+
+## obtaining correlation statistics for the dataset using modelsummary
+final_dataset |>
+  datasummary_correlation()
+```
+
+From the above table \@ref(table: summary 2) we can see that the variables with the strongest correlation with assaults in 2020 are number of female covid cases, number of male covid cases, number of covid cases for age groups less than or equal to 19, 20 to 29, 30 to 39 and 40 to 49.
+It makes for these variables to relate to number of assaults in a neighbourhood since they are the most important ones since age and gender are incredibly important in such situations when an assault is occurring.
+We also see that age groups and the number of male and female covid cases are strongly related.
+
+```{r figure 1, echo=FALSE, fig.cap="Assualts in 2020 compared to assaults in 2019"}
+## constructing a ggplot for assaults in 2019 and assaults in 2020 and using colors to create a legend
+colors<-c("Assaults 2019"="red", "Assaults 2020"="blue")
+ggplot()+
+geom_point(data=final_dataset, mapping=aes(x=Neighbourhood, y=Assault_2019,color="Assaults 2019"),size=1) + 
+geom_point(data=final_dataset, aes(x=Neighbourhood, y=Assault_2020,color="Assaults 2020"),size=1)+
+  labs(x="Neighbourhoods",
+       y="Assaults 2019, Assaults 2020",
+       title="Figure 1: Assualts in 2020 compared to assaults in 2019")
+```
+
+Fig \@ref(fig:figure 1) shows us how the number of assaults in 2019 compare to the number of assaults in 2020 per neighbourhood.
+There does not appear to be a significant difference between the values for each neighbourhood, however, it does seem like most of the assaults in 2020 were less than the ones in 2019 which again might imply that covid does affect number of assaults in a neighbourhood.
+
+## Covid cases by neighbourhood
+
+```{r figure 2, echo=FALSE, fig.cap="Covid cases for each neighborhood in 2020"}
+## constructing a ggplot to see how covid cases vary by neighbourhood
+final_dataset |>
+  ggplot(mapping=aes(x=Neighbourhood,y=Total_CASES, color="red"))+
+  geom_point()+
+  labs(x= "Neighbourhoods",
+       y="Total number of covid cases in the neighbourhood",
+       title="Figure 2: Covid cases for each neighbourhood")
+```
+
+As we can see from figure \@ref(fig:figure 2), the number of covid cases per neighbourhood varies and are not similar to each other.
+So there is variation which means there's a possibility the number of assaults might also vary with the covid cases.
 
@@
 
-\begin{equation}
-Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}  (\#eq:bayes)
-\end{equation}
+The model I will be using for this paper is:
 
-Equation \@ref(eq:bayes) seems useful, eh?
+$$ y = \beta_0 +\beta_1*x_1+\beta_2*x_2 +\beta_3*x_3 +\beta_4*x_4 $$
 
-Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR].
-We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.
+y= number of assaults in a neighbourhood
+x_1 = number of female covid cases in a neighbourhood
+x_2 = number of male covid cases in a neighbourhood
+x_3 = number of covid cases between ages 30 and 39
+x_4 = number of covid cases between ages 40 and 49
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+Here, the number of assaults is the explanatory variable and the number of female covid cases, number of make cases, number between 30 and 39 and number between 40 and 49 are the independent variables.
+
+The reason I have chosen these 4 variables is because from the correlation table \@ref(table: summary 2) in the data section, these variables have some of the strongest correlation with the number of assaults in 2020.
+
+Additionally, most crimes would not be committed by people older than 50 and would normally be committed by those in the range of 30 to 49 as those who are younger than 30 are still trying to figure out their lives and might not turn to crime yet and those older than 50 are middle aged and old people who would not normally commit crimes.
+Additionally, gender is one of the most important factors in today's world and is an important variable to include.
+It is always important to take gender into account as it can often be a confounding variable if you do not include it in the model.
+I have excluded the transgender and unknown variable from the model because those entries were far less than the number of entries for male and female.
 
 # Results
+
+Using modelsummary [@modelsummary] and AIC selection process, I was able to obtain my model and the estimates for the model such as intercept, R^2, estimates for independent variables etc which are included in the table \ref(table: model estimates).
+
+The AIC process and the code for the estimates are located in tables \ref(model).
+By the process of forwards AIC, I was able to obtain the final model by removing variables from the model to see if it reduced the AIC value.
+Once I obtained the lowest AIC value I could obtain, I had obtained my final model.
+
+## Table of estimates
+
+```{r model estimates, echo=FALSE}
+## creating 2 columns and then merging them to get the dataset
+Terms<-c("Intercept", "Female_covid_cases estimate", "Male_covid_cases estimate","30 to 39 estimate","40 to 49 estimate", "Number of observation", "R^2", "R^2 adjusted", "AIC", "BIC", "Log.Lik", "F")
+values<-c(47.418, -0.858, 1.059, 13.834,-1.964, 140 ,0.508, 0.493, 1666.5, 1684.1, -827.243, 34.810)
+data.frame(Terms, values)
+```
+## Final model
+
+So, using the estimates obtained, the model we have is:
+
+$$ y = 47.418	-0.858*x_1+1.059*x_2 +13.834*x_3 -1.964*x_4 $$
+
+This shows that there appears to be more of a positice relation between number of assaults and covid cases in a neighbourhood.
+When there is a unit increase in number of females who have covid, there is 0.858 decrease in the number of assaults in that neighbourhood.
+However, when there is a unit increase in the number of males with covid, there is a 1.059 increase in the number of assaults in the neighbourhood.
+When there is an unit increase in the number of people between 30 and 39 who get covid, there is a 13.834 increase in the number of assaults.
+However, when there is an unit increase in the number of people between 40 and 49 who get covid, there is only a 1.964 decrease in the number of assaults.
+
+## final model graphs
+
+```{r figure 3, echo=FALSE, fig.cap="Relationship between Covid and Assault rates in 2020"}
+final_dataset |>
+  ggplot(mapping = aes(x = Total_CASES, y = Assault_2020,color=""))+
+  geom_point()+
+  labs(x="Total covid cases in a neighbourhood",
+       y="Assaults in 2020",
+       title="Figure 3: Relation between covid cases and assaults")
+```
+
+From figure 3 \@ref(fig:figure 3), we can see that there appears to be a sort of positive relation between the total number of cases in a neighbourhood and the number of assaults in that neighbourhood for the year 2020.
+Except for a few outliers, it seems that when cases are low, assaults are low and when cases increase, assaults also increase.
+
+```{r figure 4, echo=FALSE}
+final_dataset |>
+  ggplot(aes(x = FEMALE_Covid_CASES, y = Assault_2020)) +
+  geom_point() +
+  geom_smooth(method = "lm",
+              color = "black", 
+              linetype = "dashed",
+              formula = 'y ~ x') +
+  labs(x = "FEMALE_Covid_CASES",
+       y = "Assaults in 2020",
+       title = "Figure 4: Relation between number of assaults in 2020 and the number of female covid cases") +
+  theme_classic() +
+  scale_color_brewer(palette = "Set1")
+```
+
+```{r figure 5, echo=FALSE}
+final_dataset |>
+  ggplot(aes(x = MALE_Covid_CASES, y = Assault_2020)) +
+  geom_point() +
+  geom_smooth(method = "lm",
+              color = "black", 
+              linetype = "dashed",
+              formula = 'y ~ x') +
+  labs(x = "MALE_Covid_CASES",
+       y = "Assaults in 2020",
+       title = "Figure 5: Relation between number of assaults in 2020 and number of male covid cases") +
+  theme_classic() +
+  scale_color_brewer(palette = "Set1")
+```
+
+```{r figure 6, echo=FALSE}
+final_dataset |>
+  ggplot(aes(x =`30 to 39`, y = Assault_2020)) +
+  geom_point() +
+  geom_smooth(method = "lm",
+              color = "black", 
+              linetype = "dashed",
+              formula = 'y ~ x') +
+  labs(x = "30 to 39",
+       y = "Assaults in 2020",
+       title = "Figure 6: Relation between number of assaults in 2020 and number of covid cases for age group 30 to 39") +
+  theme_classic() +
+  scale_color_brewer(palette = "Set1")
+```
+
+```{r figure 7, echo=FALSE}
+final_dataset |>
+  ggplot(aes(x =`40 to 49`, y = Assault_2020)) +
+  geom_point() +
+  geom_smooth(method = "lm",
+              color = "black", 
+              linetype = "dashed",
+              formula = 'y ~ x') +
+  labs(x = "40 to 49",
+       y = "Assaults in 2020",
+       title = "Figure 7: Relation between number of assaults in 2020 and number of covid cases for age group 40 to 49") +
+  theme_classic() +
+  scale_color_brewer(palette = "Set1")
+```
+
+```{r figure 8, echo=FALSE}
+final_dataset |>
+  ggplot(aes(x = FEMALE_Covid_CASES+MALE_Covid_CASES+`30 to 39`+`40 to 49`, y = Assault_2020)) +
+  geom_point() +
+  geom_smooth(method = "lm",
+              color = "black", 
+              linetype = "dashed",
+              formula = 'y ~ x') +
+  labs(x = "FEMALE_Covid_CASES+MALE_Covid_CASES+30 to 39+40 to 49",
+       y = "Assaults in 2020",
+       title = "Figure 8: Relation between number of assaults in 2020 and the final model obtained") +
+  theme_classic() +
+  scale_color_brewer(palette = "Set1")
+```
+
+Figure 4,5,6,7 and 8 all show that there seems to be a positive relation between the number of covid cases in a neighbourhood and the number of assaults in that neighbourhood.
+The strongest positive relation is shown to be between the number of assaults in a neighbourhood and the age group 30 to 39 since figure 6 has the steepest slope.
+
+```{r figure 5, echo=FALSE}
+par(mfrow = c(2,2))
+model<-
+  lm(Assault_2020 ~FEMALE_Covid_CASES+MALE_Covid_CASES+`30 to 39`+`40 to 49`,
+     data = final_dataset)
+plot(model)
+```
+
+We can analyse the goodness of fit for the model from the above 4 plots \ref(fig:figure 5).
+From the residuals vs fitted plot, we see that linearity for the model does hold and that the independent variables do enter in a linear way.
+However, from the spread of the residuals we see that variance of our residuals is not constant which means that the model has heteroscedasticity.
+From the Normal Q-Q plot, we see that our data is normally distributed since the points mostly lie on the dotted line except for a few outliers.
+
+## Conclusion of results
+
+The results obtained from our model and analysis of the dataset seem to imply that the number of covid cases in a neighbourhood had a positive relationship with the number of assaults in that same neighbourhood in the year 2020.
+The final model we obtained is:
+$$ y = 47.418	-0.858*x_1+1.059*x_2 +13.834*x_3 -1.964*x_4 $$
+where
+
+y= number of assaults in a neighbourhood
+x_1 = number of female covid cases in a neighbourhood
+x_2 = number of male covid cases in a neighbourhood
+x_3 = number of covid cases between ages 30 and 39
+x_4 = number of covid cases between ages 40 and 49
 
@@
 
-# Additional details
+# Appendix A
+
+## AIC selection
+
+For this paper, AIC selection using modelsummary was used to obtain the final model. Several models were constructed and then using modelsummary, the model with the lowest AIC value was selected.
+
+```{r, echo=TRUE}
+## the first model constructed using lm with no variables
+first_model <-
+  lm(Assault_2020 ~1,
+     data = final_dataset)
+modelsummary(first_model)
+```
+
+The AIC obtained from this model is 1757.7, so we will add a variable to see if we can reduce it.
+
+```{r, echo=TRUE}
+second_model<-
+  lm(Assault_2020 ~FEMALE_Covid_CASES,data = final_dataset)
+modelsummary(second_model)
+```
+
+The AIC obtained from this model is 1726.5 which is less so we can keep the female variable.
+
+```{r, echo=TRUE}
+third_model<-
+  lm(Assault_2020 ~FEMALE_Covid_CASES+MALE_Covid_CASES,data = final_dataset)
+modelsummary(third_model)
+```
+
+AIC obtained is 1698.6 so we can keep going.
+
+```{r, echo=TRUE}
+fourth_model<-
+  lm(Assault_2020 ~FEMALE_Covid_CASES+MALE_Covid_CASES+transgender_covid_cases,data = final_dataset)
+modelsummary(fourth_model)
+```
+
+The AIC obtained is 1700.5 so we don't include transgender covid cases
+
+```{r, echo=TRUE}
+fifth_model<-
+  lm(Assault_2020 ~FEMALE_Covid_CASES+MALE_Covid_CASES+`20 to 29`,data = final_dataset)
+modelsummary(fifth_model)
+```
+
+We keep the 20 to 29 variable since the AIC is lower.
+
+```{r, echo=TRUE}
+sixth_model<-
+  lm(Assault_2020 ~FEMALE_Covid_CASES+MALE_Covid_CASES+`20 to 29`+`30 to 39`,data = final_dataset)
+modelsummary(sixth_model)
+```
+
+So we move on with this model.
+
+```{r, echo=TRUE}
+seventh_model<-
+  lm(Assault_2020 ~FEMALE_Covid_CASES+MALE_Covid_CASES+`20 to 29`+`30 to 39`+`40 to 49`,data = final_dataset)
+modelsummary(seventh_model)
+```
+
+We proceed on with this model and will try to remove the 20 to 29 to see if we get a lower AIC.
+
+```{r, echo=TRUE}
+eight_model<-
+  lm(Assault_2020 ~FEMALE_Covid_CASES+MALE_Covid_CASES+`30 to 39`+`40 to 49`,data = final_dataset)
+modelsummary(eight_model)
+```
+
+The AIC is even lower so we remove 20 to 29. We try with another variable
+
+```{r, echo=TRUE}
+seventh_model<-
+  lm(Assault_2020 ~FEMALE_Covid_CASES+MALE_Covid_CASES+`40 to 49`,data = final_dataset)
+modelsummary(seventh_model)
+```
+
+This time the AIC is higher.
+
+```{r, echo=TRUE}
+seventh_model<-
+  lm(Assault_2020 ~FEMALE_Covid_CASES+MALE_Covid_CASES+`30 to 39`,data = final_dataset)
+modelsummary(seventh_model)
+```
+
+The AIC is again higher than the AIC value for the eight model. So the eight model is our final model for this paper.
+
+Assault_2020 ~FEMALE_Covid_CASES+MALE_Covid_CASES+`30 to 39`+`40 to 49`
+
+These are the estimates obtained for the final model using modelsummary.
+
+```{r model, echo=TRUE}
+final_model<-
+  lm(Assault_2020 ~FEMALE_Covid_CASES+MALE_Covid_CASES+`30 to 39`+`40 to 49`,
+     data = final_dataset)
+modelsummary(final_model)
+```
+
+# Appendix B
+
+## Covid Cases
+
+This is a preview of the covid_cases dataset obtained from opendatatoronto [@citeopendatatoronto]
+
+```{r dataset1, echo=FALSE}
+covid_cases
+```
+
+## Neighbourhood crime rates
+
+This is a preview of the neighbourhood crimes dataset obtained from opendatatoronto [@citeopendatatoronto]
+
+```{r dataset2, echo=FALSE}
+neighbourhood_crime_rates
+```
+
+## Appendix C
 
@@
 # References
+
+<div id = "refs"></div>

### 2022-04-27 outputs/paper/paper.Rmd
most of structure is complete

@@
 ---
-title: "Modelling Covid's effect on Assaults in Toronto"
-subtitle: "A statistical analysis of how the rise of Covid in 2020 affected assault rates in various downtown neighborhoods in Toronto"
+title: "Does Covid encourage assaults?"
+subtitle: "A statistical analysis of how the rise of Covid in 2020 affected assault rates in various neighborhoods in Toronto"
 author: 
@@
 The data was extracted from the provincial Case & Contact Management System and the variables in the dataset include the neighbourhood the case was reported in, the date it was reported, the client gender, Age group and the classification of the case, whether it was probable or improbable.
-The full dataset /@ref(table:dataset1) can be found in Appendix B.
+The full dataset \@ref(table:dataset1) can be found in Appendix B.
 
@@
 We are only interested in the number of assaults for 2019 and 2020.
-The full dataset /@ref(table:dataset2) can be found in Appendix B.
+The full dataset \@ref(table:dataset2) can be found in Appendix B.
 
-The datasets were extracted using R [@citeR] and R packages opendatatoronto [@citeopendatatoronto], tidyverse [@tidyverse], dpylr [@dpylr] and knitr [@knitr].
+The datasets were extracted using R [@citeR] and R packages opendatatoronto [@citeopendatatoronto], tidyverse [@tidyverse], dplyr [@dplyr] and knitr [@knitr].
 The data was then processed, manipulated and analysed for the purposes of this paper.
@@
 
-The final dataset /@ref(table:dataset_final) consists of 140 observations and 13 variables:
+The final dataset \@ref(table:dataset_final) consists of 140 observations and 13 variables:
 Neighbourhood: Name of the neighbourhood
@@
 
-## final model graphs
+## Final model graphs
 
@@
 
-```{r figure 5, echo=FALSE}
+```{r figure 9, echo=FALSE}
 par(mfrow = c(2,2))
@@
 where 
-
 y= number of assaults in a neighbourhood
@@
 <div id = "refs"></div>
+
+[@thereferencecanbewhatever]
+
+[@ggplot2]
+
+[@dplyr]
+
+[@tidyr]
+
+[@janitor]
+
+[@reshape2]
+
+[@here]
+
+[@tinytex]
+
+[@readr]
+
+[@metaopendata]

### 2022-04-27 scripts/final-data-prep.R
commiting changes to scripts and readme

@@ -1,0 +1,63 @@
+
+# Covid cases modification
+
+
+## extracting the dataset
+options(readr.show_col_types = FALSE)
+covid_cases<-
+  read_csv(here::here("inputs/data/covid_cases.csv"))
+covid_cases<-
+  covid_cases |>
+  mutate(year=format(covid_cases$`Episode Date`,"%Y"))
+
+## isolating all cases from 2020
+covid_cases2020<-
+  covid_cases |>
+  filter(year == 2020)
+covid_cases_gender<-covid_cases2020 %>% group_by(`Neighbourhood Name`, `Client Gender`) %>% count()
+covid_cases_age<-covid_cases2020 %>% group_by(`Neighbourhood Name`, `Age Group`) %>% count()
+## creating required variables such as Female covid cases and male covid cases
+covid_cases_gender<-
+  covid_cases_gender |>
+  pivot_wider(names_from = `Client Gender`,values_fill =0,values_from = n)
+## creating required variables for age groups
+covid_cases_age<-
+  covid_cases_age |>
+  pivot_wider(names_from = `Age Group`,values_fill =0,values_from = n)
+## mergind the two datasets together to obtain final dataset
+covid_cases_final<-merge(covid_cases_gender,covid_cases_age)
+covid_cases_final<-
+  covid_cases_final |>
+  mutate(total_cases=FEMALE+MALE+TRANSGENDER+UNKNOWN+OTHER)
+covid_cases_final<-covid_cases_final |>
+  select(`Neighbourhood Name`, FEMALE,MALE,TRANSGENDER,UNKNOWN,OTHER, `19 and younger`,`20 to 29 Years`,`30 to 39 Years`,`40 to 49 Years`,`50 to 59 Years`,`60 to 69 Years`,`70 to 79 Years`,`80 to 89 Years`,`90 and older`, total_cases)
+colnames(covid_cases_final)[1]<-"Neighbourhood"
+colnames(covid_cases_final)[2]<-"FEMALE_Covid_Cases"
+colnames(covid_cases_final)[3]<-"MALE_Covid_Cases"
+covid_cases_final<-na.omit(covid_cases_final)
+
+
+# Neighbourhood cases modification
+
+
+## extracting the dataset
+neighbourhood_crime_rates<-
+  read_csv(here::here("inputs/data/neighbourhood_crime_rates.csv"))
+neighbourhood_crime_rates_assaults<- 
+  neighbourhood_crime_rates |>
[…]

### 2022-04-27 scripts/final-data-prep.R
final changes

@@ -34,7 +34,7 @@ covid_cases_final<-covid_cases_final |>
 colnames(covid_cases_final)[1]<-"Neighbourhood"
 colnames(covid_cases_final)[2]<-"FEMALE_Covid_Cases"
 colnames(covid_cases_final)[3]<-"MALE_Covid_Cases"
-covid_cases_final<-na.omit(covid_cases_final)
+covid_cases_final<-na.omit(covid_cases_final) ## cleaning one na case
 
 
 # Neighbourhood cases modification
@@ -46,7 +46,6 @@ neighbourhood_crime_rates<-
 neighbourhood_crime_rates_assaults<- 
   neighbourhood_crime_rates |>
   select(Neighbourhood,
-         Assault_2019,
          Assault_2020)
 neighbourhood_crime_rates_assaults<-arrange(neighbourhood_crime_rates_assaults,Neighbourhood)
 

### 2022-04-27 outputs/paper/paper.Rmd
commiting final changes and completed assignment

--- 2022-04-27 outputs/paper/paper.Rmd
@@
 ---
-title: "Modelling Covid's effect on Assaults in Toronto"
-subtitle: "A statistical analysis of how the rise of Covid in 2020 affected assault rates in various downtown neighborhoods in Toronto"
+title: "Did covid encourage assaults in Toronto neighbourhoods in 2020?"
+subtitle: "A statistical analysis of how the rise of Covid in 2020 affected assault rates in various neighborhoods in Toronto"
 author: 
@@
 date: "`r format(Sys.time(), '%d %B %Y')`"
-abstract: "This paper presents an analysis on the effect of Covid on assault rates in downtown Toronto neighbourhoods in 2020.
-The paper uses datasets obtained from opendatatoronto to show how Covid had a negative effect on crime rates.
+abstract: "This paper presents an analysis on the effect of Covid on assault rates in Toronto neighbourhoods in 2020.
+The paper uses datasets obtained from opendatatoronto to show how covid has a positive relation with assaults in a neighbourhood.
 The analysis consists of using the statistical programming language R to model the data using regression and constructing various plots and graphs to visually represent the data.
-THe results obtained will allow us to better predict people's behavior.
+The results obtained will allow us to better predict people's behavior.
 With the better prediction of people's behaviour, the police can better manage the public order and government will be able to make reasonable and effective policy to reduce the crime rate.\\par\\textbf {Keywords:} Covid, Assaults, Crime,Toronto neighbourhoods, residents, open data toronto"
@@
 library(modelsummary)
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, out.width = "65%", out.height = "65%")
 ```
 
-# 1 Introduction
+# Introduction
 
@@
 Covid had a huge impact on the global economy and had restricted people's actions and movements.
-However, this paper wants to analyse how Covid has affected one of the most common types of crime, assaults.
-Assaults are often physical and in a time where physical contact is discouraged, it seems contradictory to assault someone as there is the risk you might catch covid.
+However, this paper wants to analyse how the rise of Covid has affected one of the most common types of crime, assaults, which are often physical in nature.
 The question this paper seeks to answer is how the Covid pandemic had affected assault rates in various neighbourhoods in Toronto for the year 2020.
 
-In this paper, we use data on Covid cases and crime statistics in Toronto neighbourhoods for the year 2020 to see how the number of Covid cases in a neighbourhood had affected assualts in that neighbourhood and what factors affected the assault rate the most.
-We constructed a model using R in order to see how Covid can affect the assault rate in a neighborhood.
-We also use the dataset to construct various histograms, bar plots and summary statistics about the dataset to improve the quality of our analysis.
-The results obtained seem to imply that the Covid pandemic had a negative effect on assaults in 2020 as crime rates went down in most downtown neighborhoods.
-The results obtained contribute to our understanding of how the presence of a common threat that affects everyone such as a pandemic in this case can discourage people from criminal actions.
+In this paper, we use data on Covid cases and crime statistics in Toronto neighbourhoods for the year 2020 to see how the number of Covid cases in a neighbourhood had affected assaults in 2020 and what factors affected the assault rate the most.
+We constructed a linear model using R [@citeR] to model covid's effect on assaults.
+We also used the dataset to construct various plots and summary statistics about the dataset to improve the quality of our analysis.
+The results obtained seem to imply that the Covid pandemic had a positive effect on assaults in 2020 as there is a positive relation between number of covid cases in a neighbourhood and number of assaults in Toronto.
+
+The results obtained from this paper help contribute to our understanding of how the presence of a common threat that affects everyone such as a pandemic in this case can discourage people from criminal actions.
 It will allow us to improve our predictions of a person's behavior during such situations.
-Better predictions of people's behaviors will allow the police to better manage the public order and government to make reasonable and effective policy to reduce the crime rate
+Better predictions of people's behaviors will allow the police to better manage public order and government to make reasonable and effective policy to reduce the crime rate during such situations.
 
 The paper first presents an overview of the datasets and the variables we will be using for this study in the data section.
-The model section presents the regression steps taken to generate the model that shows how Covid affects crime.
-The results section discusses the various conclusions we can draw from the data and the model and finally the discussion section discusses the strengths and weaknesses of the paper.
+The model section presents the variables chosen for our model and the justification for these variables.
+The results section discusses the steps taken to generate estimates for our final model and the various conclusions we can draw from  both the data and the model and finally the discussion section discusses the strengths and weaknesses of the paper.
 Any additional details are included in the appendix.
+An enhancement is included in appendix C.
 
-# 2 Data
+# Data
 
-## 2.1 Data Source and collection
+## Data Source and collection
 
-For the dataset, we obtain data using opendatatoronto on crime related statistics, Covid related data pertaining to Toronto neighborhoods.
+The relevant datasets were obtained using opendatatoronto and show statistics related to Covid cases and to crime figures.
+The first dataset is about Covid cases in Toronto and records all Covid cases reported since January 2020.
+The dataset includes all cases related to outbreaks and sporadic cases, those that occur in the community.
+The data was extracted from the provincial Case & Contact Management System and the variables in the dataset include the neighbourhood the case was reported in, the date it was reported, the client gender, Age group and the classification of the case, whether it was probable or improbable.
+The full dataset \@ref(table: dataset1) can be found in Appendix B.
 
-## 2.2 Data Modification
+The second dataset used for this paper was the dataset related to crime statistics in Toronto neighbourhoods.
+The dataset contains a list of crime related statistics by neighbourhood and includes figures such as number of assaults, number of car thefts, number of shootings, etc.
+for the years 2014 to 2020.
+We are only interested in the number of assaults for 2020.
+The full dataset \@ref(table: dataset2) can be found in Appendix B.
 
-## 2.3 Data Tables
+The datasets were extracted using R [@citeR] and R packages opendatatoronto [@citeopendatatoronto], tidyverse [@tidyverse], dplyr [@dplyr] and knitr [@knitr].
+The data was then processed, manipulated and analysed for the purposes of this paper.
 
-```{r table 1, echo=FALSE}
-options(readr.show_col_types = FALSE)
-covid_cases<-
-  read_csv(here::here("inputs/data/covid_cases.csv"))
-covid_cases<-
-  covid_cases |>
-  mutate(year=format(covid_cases$`Episode Date`,"%Y"))
-covid_cases2020<-
-  covid_cases |>
-  filter(year == 2020)
-covid_cases_gender<-covid_cases2020 %>% group_by(`Neighbourhood Name`, `Client Gender`) %>% count()
-covid_cases_gender<-
-  covid_cases_gender |>
-  pivot_wider(names_from = `Client Gender`,values_fill =0,values_from = n)
-covid_cases_outbreak<-covid_cases2020 %>% group_by(`Neighbourhood Name`,`Outbreak Associated`) %>% count()
-covid_cases_outbreak<-
-  covid_cases_outbreak |>
-  pivot_wider(names_from = `Outbreak Associated`,values_fill =0,values_from = n)
-covid_cases_final<-merge(covid_cases_gender,covid_cases_outbreak)
-covid_cases_final<-
-  covid_cases_final |>
-  mutate(total_cases=`Outbreak Associated`+Sporadic)
-colnames(covid_cases_final)[1]<-"Neighbourhood"
-colnames(covid_cases_final)[2]<-"FEMALE_Covid_Cases"
-colnames(covid_cases_final)[3]<-"MALE_Covid_Cases"
-covid_cases_final<-na.omit(covid_cases_final)
-```
+## Data Modification
 
-```{r table 2, echo=FALSE}
-neighbourhood_crime_rates<-
-  read_csv(here::here("inputs/data/neighbourhood_crime_rates.csv"))
-neighbourhood_crime_rates_assaults<- 
-  neighbourhood_crime_rates |>
-  select(Neighbourhood,
-         Assault_2019,
-         Assault_2020)
-neighbourhood_crime_rates_assaults<-arrange(neighbourhood_crime_rates_assaults,Neighbourhood)
-```
+For the purposes of this paper, we are interested only in covid cases that were reported during 2020 since that was when the pandemic first started.
+So the only cases selected were those reported during 2020.
+We then focused on the variables Client Gender and the age group of the infected.
+These variables are the ones that pertain the most to the total number of covid cases in a neighbourhood and so are the ones we will be focusing on for this dataset as they also more important than the others.
+New variables using the dataset were created such as number of female covid cases, number of male covid cases, number of transgender covid cases, number of other genders and unknown genders covid cases and variables related to the various age groups, less than 19, between 20 to 29, 30 to 39, etc.
+We also created a new variable where we counted the total number of Covid cases in a neighbourhood.
+However, we will be excluding all age groups greater than or equal to 50 since it is unlikely that anyone of that age will be assaulting people as they are mostly middle aged or old people.
+All NA cases were also deleted in order to improve the analysis.
 
-## Obtaining the final dataset.
+For the crime dataset, we only focused on the neighbourhoods and the number of assaults for the year 2020 since these pertain directly to my paper.
+We then joined the two datasets together and obtained the final_dataset.
+All 140 neighbourhoods are included in this dataset and it is the dataset we will be using for the rest of this paper.
 
-```{r table 3, echo=FALSE, fig.cap="Final Dataset"}
-covid_cases_final<-arrange(covid_cases_final,Neighbourhood)
-neighbourhood_crime_rates_assaults<-arrange(neighbourhood_crime_rates_assaults,Neighbourhood)
+The code for the dataset modifications are included in scripts in final-data-prep.R.
+
+## Dataset for this paper
+
+```{r dataset_final, echo=FALSE}
 final_dataset<-
-  neighbourhood_crime_rates_assaults |>
-  mutate(FEMALE_Covid_CASES=covid_cases_final$FEMALE_Covid_Cases,MALE_Covid_CASES=covid_cases_final$MALE_Covid_Cases,transgender_covid_cases=covid_cases_final$TRANSGENDER,unknown_covid_cases= covid_cases_final$UNKNOWN,other_covid_cases=covid_cases_final$OTHER, Outbreak_Associated=covid_cases_final$`Outbreak Associated`,Sporadic=covid_cases_final$Sporadic,Total_CASES=covid_cases_final$total_cases)
+  read_csv(here::here("inputs/data/final_dataset.csv"))
 final_dataset
@@
 
-## 2.4 Data Visualization
+The final dataset \@ref(table: dataset_final) consists of 140 observations and 12 variables:
+Neighbourhood: Name of the neighbourhood,
+Assault_2020: number of assaults in that neighbourhood for the year 2020,
+FEMALE_Covid_CASES: Number of reported covid cases in that neighbourhood for females,
+MALE_Covid_CASES: Number of reported covid cases in that neighbourhood for males,
+transgender_Covid_CASES: Number of reported covid cases in that neighbourhood for transgenders,
+unknown_covid_cases: Number of reported covid cases in that neighbourhood for people of unknown genders,
+other_covid_cases: Number of reported covid cases in that neighbourhood for people of other genders,
+<=19: number of reported covid cases for people less than 19 years old,
+20 to 29: number of reported covid cases for people between 20 and 29,
+30 to 39: number of reported covid cases for people between 30 and 39,
+40 to 49: number of reported covid cases for people between 40 and 49,
+Total_CASES: total number of covid cases in a neighbourhood.
+
+These 12 variables are the ones that are most pertinent to the topic of this paper. And in this dataset, the number of assaults in a neighbourhood will the response variable for this paper.
+
+## Summary statistics
 
 ```{r summary 1, echo=FALSE}
-final_dataset |>
+
+## obtaining summary statistics for the dataset using modelsummary
+summary<-final_dataset |>
   datasummary_skim()
+summary
 ```
 
+The above table \@ref(table: summary 1) shows summary statistics for the dataset.
+As you can see, on average, the number of assaults per neighbourhood in Toronto was 127.5 for 2020 with a standard deviation of 127.5 which means that there was a variation in assaults per neighboruhood.
+Furthermore, the number of covid cases on average in a neighbourhood is 225.4 with a standard deviation of 206.3 so each neighbourhood has a varying amount of covid cases.
+So it is possible there is a relation between number of assault cases in a neighbourhood and number of covid cases.
+
+## Correlation statistics for the final dataset
+
 ```{r summary 2,echo=FALSE}
+
+## obtaining correlation statistics for the dataset using modelsummary
 final_dataset |>
@@
 
-```{r figure 1, echo=FALSE, fig.cap="Assualts in 2020 compared to assaults in 2019"}
-colors<-c("Assaults 2019"="red", "Assaults 2020"="blue")
+From the above table \@ref(table: summary 2) we can see that the variables with the strongest correlation with assaults in 2020 are number of female covid cases, number of male covid cases, number of covid cases for age groups 20 to 29, 30 to 39 and 40 to 49.
+It makes for these variables to relate to number of assaults in a neighbourhood since they are the most important ones since age and gender are incredibly important in such situations when an assault is occurring.
+We also see that age groups and the number of male and female covid cases are strongly related.
+
+## Assault cases by neighbourhood
+
+```{r figure 1, echo=FALSE,fig.dim = c(8, 6), fig.cap="Assualts in 2020 by neighbourhood"}
+## constructing a ggplot for assaults in 2020
 ggplot()+
-geom_point(data=final_dataset, mapping=aes(x=Neighbourhood, y=Assault_2019,color="Assaults 2019"),size=1) + 
-geom_point(data=final_dataset, aes(x=Neighbourhood, y=Assault_2020,color="Assaults 2020"),size=1)+
+geom_point(data=final_dataset, aes(x=Neighbourhood, y=Assault_2020,color="red"),size=1)+
   labs(x="Neighbourhoods",
-       y="Assaults 2019, Assaults 2020",
-       title="Assualts in 2020 compared to assaults in 2019")
+       y="Assaults 2020",
+       title="Figure 1: Assualts in 2020 by neighbourhood")
 ```
 
-```{r figure 2, echo=FALSE, fig.cap="Covid cases for each neighborhood in 2020"}
-colors<-c("Female"="red", "Male"="blue", "Transgender"="purple","Unknown"="green","Other"="yellow")
-ggplot()+
-geom_point(data=final_dataset, mapping=aes(x=Neighbourhood, y=FEMALE_Covid_CASES,color="Female"),size=1) + 
-geom_point(data=final_dataset, aes(x=Neighbourhood, y=MALE_Covid_CASES,color="Male"),size=1)+
-  geom_point(data=final_dataset, aes(x=Neighbourhood, y=transgender_covid_cases,color="Transgender"),size=1)+
-  geom_point(data=final_dataset, aes(x=Neighbourhood, y=unknown_covid_cases,color="Unknown"),size=1)+
-  geom_point(data=final_dataset, aes(x=Neighbourhood, y=other_covid_cases,color="Other"),size=1)+
-  labs(x="Neighbourhoods",
-       y="Female, Male, Transgender, Unknown, Other",
-       title="Covid cases by gender and neighborhood")
+Fig \@ref(fig: figure 1) shows us the number of assaults in 2020 per neighbourhood.
+There does seem to be some slight variation in the number of assaults per neighbourhood with some high outliers which match the results shown in the summary table.
+
+## Covid cases by neighbourhood
+
+```{r figure 2, echo=FALSE,fig.dim = c(8, 6), fig.cap="Covid cases for each neighborhood in 2020"}
+## constructing a ggplot to see how covid cases vary by neighbourhood
+final_dataset |>
+  ggplot(mapping=aes(x=Neighbourhood,y=Total_CASES, color="red"))+
+  geom_point()+
+  labs(x= "Neighbourhoods",
+       y="Total number of covid cases in the neighbourhood",
+       title="Figure 2: Covid cases for each neighbourhood")
 ```
+
+As we can see from figure \@ref(fig: figure 2), the number of covid cases per neighbourhood varies and are not similar to each other.
+Additionally, the variation pattern in the number of covid cases seems to be similar to the pattern in the number of assaults per neighbourhood from figure \@ref(fig: figure 1.) So, the plots support our hypothesis that there is a relation between covid cases and number of assaults in a neighbourhood.
 
@@
 
-$$ y = \beta_0 +\beta_1*x_1+\beta_2*x_2  $$
+The model we will be using for this paper is:
 
-y= number of assaults in a neighbourhood
-x_1 = number of female covid cases in a neighbourhood
-x_2 = number of male covid cases in a neighbourhood
+$$ y = \beta_0 +\beta_1*x_1+\beta_2*x_2 +\beta_3*x_3 +\beta_4*x_4 +\beta_5*x_5 $$
 
-Equation \@ref(eq:bayes) seems useful, eh?
+y= number of assaults in a neighbourhood,
+x_1 = number of female covid cases in a neighbourhood,
+x_2 = number of male covid cases in a neighbourhood,
+x_3 = number of covid cases for people between ages 20 and 29
+x_4 = number of covid cases for people between ages 30 and 39,
+x_5 = number of covid cases for people between ages 40 and 49.
 
-Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR].
-We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.
+Here, the number of assaults is the response variable and the number of female covid cases, number of male cases, number of covid cases for age group 20 to 29, number of covid cases for age group 30 to 39 and number for 40 to 49 are the explanatory variables.
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+The reason I have chosen these 5 variables is because from the correlation table \@ref(table: summary 2) in the data section, these variables have some of the strongest correlation with the number of assaults in 2020.
+
+Gender is one of the most important factors in today's world and is an important variable to include.
+It is always important to take gender into account as it can often be a confounding variable if you do not include it in the model.
+I have excluded the transgender and unknown variable from the model because those entries were far less than the number of entries for male and female.
+Since the male and female cases compose the majority of the dataset it is likely that they will have a higher effect on the response variable so we will be using these variables for the model.
+
+People older than 50 would not normally be commiting assaults so we will exclude them from the dataset and also exclude people younger than 20 since they would be teenagers and children.
+So we will only be focusing on the age groups in the range of 20 to 49 for this model.
 
@@
 
-## Table of estimates
+Using modelsummary [@modelsummary] and AIC selection process, we were able to obtain the final model and the estimates for the model such as intercept, R^2, estimates for independent variables etc which are included in the table \ref(table: model estimates).
+
+The AIC process and the code for the estimates are located in tables \ref(eq: model).
+By the process of forwards AIC, we were able to obtain the final model by removing variables from the model to see if it reduced the AIC value.
+Once we obtained the lowest AIC value possible and could not add more variables, we had obtained my final model.
+
+## Model estimates
 
 ```{r model estimates, echo=FALSE}
-Terms<-c("Intercept", "Female_covid_cases estimate", "Male_covid_cases estimate", "Number of observation", "R^2", "R^2 adjusted", "AIC", "BIC", "Log.Lik", "F")
-values<-c(58.410, -1.174, 2.029, 140, 0.363, 0.354, 1698.6, 1710.3, -845.280, 39.044)
+## creating 2 columns and then merging them to get the dataset
+Terms<-c("Intercept", "Female_covid_cases estimate", "Male_covid_cases estimate","20 to 29 estimate", "30 to 39 estimate","40 to 49 estimate", "Number of observation", "R^2", "R^2 adjusted", "AIC", "BIC", "Log.Lik", "F")
+values<-c(64.112, -0.918, -0.343,2.06, 4.37,-1.573, 140 ,0.397, 0.375, 1696.9, 1717.5, -841.440, 17.648)
 data.frame(Terms, values)
@@
 
-$$ y = 58.410	 -1.174*x_1+2.029*x_2  $$
+## Final model
 
-```{r figure , echo=FALSE, fig.cap="Relationship between Covid and Assault rates in 2020"}
+So, using the estimates obtained, the model we have is:
+
+$$ y = 64.112	-0.918*x_1+-0.343*x_2 +2.060*x_3 +4.370*x_4 -1.573*x_5 $$
+
+The estimates for our model show us that there seems to be a positive relationship between the number of covid cases in a neighbourhood and the number of assaults.
+When there is an increase in the number of male or female cases in a neighbourhood, then there is a small decrease in the number of assaults.
+However, if the age group affect was 20 to 29 or 30 to 39, the positive increase in the number of assaults is greater than the decrease due to x1 and x2.
+Only when the age group is of 40 to 49 does the assaults actually decrease but since two of the age groups have a positive effect on number of assaults, we can say that the number of cases of covid in a neighbourhood will have a positive effect on the number of assaults in that neighbourhood.
+
+So when number of covid cases in a neighbourhood increase, so do the number of assaults.
+From our analysis of the dataset and the model, this means that in neighbourhoods which had higher covid cases in 2020 also had a higher number of assaults for that year.
+
+## Final model graphs
+
+We will use graphs to confirm our results by modelling covid's effect on assaults in a neighbourhood.
+
+```{r figure 3, echo=FALSE,fig.dim = c(8, 6), fig.cap="Relationship between Covid and Assault rates in 2020"}
 final_dataset |>
-  ggplot(mapping = aes(x = Total_CASES, y = Assault_2020,color="")) +
-  geom_point()
+  ggplot(mapping = aes(x = Total_CASES, y = Assault_2020,color=""))+
+  geom_point()+
+  labs(x="Total covid cases in a neighbourhood",
+       y="Assaults in 2020",
+       title="Figure 3: Relation between covid cases and assaults")
 ```
 
-```{r, echo=FALSE}
+From figure 3 \@ref(fig:figure 3), we can see that there appears to be a sort of positive relation between the total number of cases in a neighbourhood and the number of assaults in that neighbourhood for the year 2020.
+Except for a few outliers, it seems that when cases are low, assaults are low and when cases increase, assaults also increase.
+
+```{r figure 4,fig.dim = c(8, 6), fig.cap="Relation between number of assaults in 2020 and the number of female covid cases", echo=FALSE}
 final_dataset |>
-  ggplot(aes(x = FEMALE_Covid_CASES+MALE_Covid_CASES, y = Assault_2020)) +
+  ggplot(aes(x = FEMALE_Covid_CASES, y = Assault_2020)) +
   geom_point() +
@@
               formula = 'y ~ x') +
-  labs(x = "FEMALE_Covid_CASES+MALE_Covid_CASES",
-       y = "total_Crime_2020") +
+  labs(x = "FEMALE_Covid_CASES",
+       y = "Assaults in 2020",
+       title = "Figure 4: Relation between number of assaults in 2020 and the number of female covid cases") +
   theme_classic() +
@@
 
-```{r}
-plot(final_model)
+```{r figure 5, fig.dim = c(8, 6), echo=FALSE}
+final_dataset |>
+  ggplot(aes(x = MALE_Covid_CASES, y = Assault_2020)) +
+  geom_point() +
+  geom_smooth(method = "lm",
+              color = "black", 
+              linetype = "dashed",
+              formula = 'y ~ x') +
+  labs(x = "MALE_Covid_CASES",
+       y = "Assaults in 2020",
+       title = "Figure 5: Relation between number of assaults in 2020 and number of male covid cases") +
+  theme_classic() +
+  scale_color_brewer(palette = "Set1")
 ```
+
+```{r figure 6,fig.dim = c(8, 6), echo=FALSE}
+final_dataset |>
+  ggplot(aes(x =`20 to 29`, y = Assault_2020)) +
+  geom_point() +
+  geom_smooth(method = "lm",
+              color = "black", 
+              linetype = "dashed",
+              formula = 'y ~ x') +
+  labs(x = "covid cases in age group 20 to 29",
+       y = "Assaults in 2020",
+       title = "Figure 6: Relation between number of assaults in 2020 and number of covid cases for age group 20 to 29") +
+  theme_classic() +
+  scale_color_brewer(palette = "Set1")
+```
+
+```{r figure 7,fig.dim = c(8, 6), echo=FALSE}
+final_dataset |>
+  ggplot(aes(x =`30 to 39`, y = Assault_2020)) +
+  geom_point() +
+  geom_smooth(method = "lm",
+              color = "black", 
+              linetype = "dashed",
+              formula = 'y ~ x') +
+  labs(x = "covid cases in age group 30 to 39",
+       y = "Assaults in 2020",
+       title = "Figure 7: Relation between number of assaults in 2020 and number of covid cases for age group 30 to 39") +
+  theme_classic() +
+  scale_color_brewer(palette = "Set1")
+```
+
+```{r figure 8,fig.dim = c(8, 6), echo=FALSE}
+final_dataset |>
+  ggplot(aes(x =`40 to 49`, y = Assault_2020)) +
+  geom_point() +
+  geom_smooth(method = "lm",
+              color = "black", 
+              linetype = "dashed",
+              formula = 'y ~ x') +
+  labs(x = "covid cases in age group 40 to 49",
+       y = "Assaults in 2020",
+       title = "Figure 8: Relation between number of assaults in 2020 and number of covid cases for age group 40 to 49") +
+  theme_classic() +
+  scale_color_brewer(palette = "Set1")
+```
+
+```{r figure 9,fig.dim = c(8, 6), echo=FALSE}
+final_dataset |>
+  ggplot(aes(x = FEMALE_Covid_CASES+MALE_Covid_CASES+`20 to 29`+`30 to 39`+`40 to 49`, y = Assault_2020)) +
+  geom_point() +
+  geom_smooth(method = "lm",
+              color = "black", 
+              linetype = "dashed",
+              formula = 'y ~ x') +
+  labs(x = "FEMALE_Covid_CASES+MALE_Covid_CASES+20 to 29+30 to 39+40 to 49",
+       y = "Assaults in 2020",
+       title = "Figure 9: Relation between number of assaults in 2020 and the final model obtained") +
+  theme_classic() +
+  scale_color_brewer(palette = "Set1")
+```
+
+Figure 4,5,6,7,8 and 9 all show that there seems to be a positive relation between the number of covid cases in a neighbourhood and the number of assaults in that neighbourhood except for a few outliers.
+When the number of covid cases in a neighbourhood are low so are the number of assaults and when the number of covid cases increase so do the number of assaults increase with it.
+
+```{r figure 10,fig.dim = c(8, 6), fig.cap="Figure 9: Goodness of fit plots", echo=FALSE}
+par(mfrow = c(2,2))
+model<-
+  lm(Assault_2020 ~FEMALE_Covid_CASES+MALE_Covid_CASES+`30 to 39`+`40 to 49`,
+     data = final_dataset)
+plot(model)
+```
+
+We can analyse the goodness of fit for the model from the above 4 plots \ref(fig:figure 10).
+From the residuals vs fitted plot, we see that linearity for the model does hold and that the independent variables do enter in a linear way.
+However, from the spread of the residuals we see that variance of our residuals is not constant which means that the model has heteroscedasticity.
+From the Normal Q-Q plot, we see that our data is normally distributed since the points mostly lie on the dotted line except for a few outliers.
+
+## Conclusion of results
+
+The results obtained from our model and analysis of the dataset seem to imply that the number of covid cases in a neighbourhood had a positive relationship with the number of assaults in that same neighbourhood in the year 2020.
+The final model we obtained is:
+
+$$ y = 64.112	-0.918*x_1+-0.343*x_2 +2.060*x_3 +4.370*x_4 -1.573*x_5 $$
+
+where 
+y= number of assaults in a neighbourhood
+x_1 = number of female covid cases in a neighbourhood
+x_2 = number of male covid cases in a neighbourhood
+x_3= number of covid cases between ages 20 and 29
+x_4 = number of covid cases between ages 30 and 39
+x_5 = number of covid cases between ages 40 and 49
+
+The number of assaults were higher in neighbourhoods with higher covid cases and were lower in neighbourhoods with lower number of covid cases.
 
@@
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+The paper analysed datasets about neighbourhood crime statistics and covid cases reported since January 2020 obtained using opendatatoronto in order to see how number of covid cases in a neighbourhood affect the number of assaults.
+Here the number of assaults in a neighbourhood was the response variable.
+The paper focused only on the data during 2020 since that was the year covid was first reported and we were interested in it's initial effects on crime.
+A linear regression model with several explanatory variables was built using the data obtained and several graphs were created based on this model.
+Using AIC we were able to build a model and obtain estimates for the various variables.
 
-## Second discussion point
+We found that the variables that affected the response variable were the number of females covid cases in a neighbourhood, number of male covid cases, number of covid cases for age group 20 to 29 and age group 30 to 39.
+We also found that the number of covid cases in a neighbourhood had a positive relation with the number of assaults in a neighbourhood.
+Neighbourhoods with higher number of covid cases also had a higher number of assaults.
+Various graphs plotted using the data confirmed these results for us as they showed a positive and linear relation between number of covid cases in a neighbourhood and number of assaults.
 
-## Third discussion point
+The results obtained supports the theory that pandemics such as covid increase the number of assaults in an area that has been severely afflicted in such situations.
+Desperation leads to panic which might cause people to inflict violence and assault others.
+It shows us that when people are placed in situations of high stress, they might panic and resort to violence especially when the stress start to increase and people around them fall ill to some affliction.
+Our model also showed us that people who were younger and were young adults were more susceptible to panic than older people
 
-## Weaknesses and next steps
+The results obtained in this paper could be used to inform police actions in the future at the beginning of situations like Covid where the entire city is impacted.
+Public order could be maintained by having police increase policing in neighbourhoods that are suffering the most to help protect citizens or have governments make better policy decisions during such situations so that people do not feel stressed and react violently.
+As an example, they could ensure better care for those in neighbourhoods that are suffering the most from the situation and could focus on younger age groups.
 
-Weaknesses and next steps should also be included.
+## Weaknesses
+
+There are a few weaknesses in this paper, the datasets used did not include factors such as race of the person with covid which could become a limitation as race might be a confounding factor.
+However, it might be unethical to try and have race in the model.
+
+Another limitation for the paper is that we were not able to obtain economic data for Toronto neighbourhoods for the year 2020.
+Economic data such as income of households or families is one of the most important variables to account for when analysing crime since income of a person often dictates whether they would turn to crime or not.
+Since, we were not able to obtain economic data for the year 2020, it is possible that our paper will be biased in its analysis.
+
+## Next steps
+
+The next steps to improve the model obtained here is to increase the data used for this model.
+Economic data for Toronto neighbourhoods should be obtained for the year 2020 in order to include these variables in the model.
+If we are able to obtain such data and include it in the future, we will be able to avoid the effect of confounding variables and will be able to obtain an unbiased model.
+
+Another limitation that could be accounted for is racial data but profiling such data might not be ethical as we will be including race in a model on crime which will lead to racial profiling.
+This paper recommends that race not be included in such models since it is unethical to model it.
+
+So, in order to fully comprehend people's behaviours in such situation more data is required in order to obtain an unbiased model that will help contribute to out understanding of people's behaviours during high stress situations.
 
@@
 
+# Appendix A
+
 ## AIC selection
-```{r, echo=FALSE}
+
+For this paper, AIC selection using modelsummary was used to obtain the final model. Several models were constructed and then using modelsummary, the model with the lowest AIC value was selected.
+
+```{r, echo=TRUE}
+## the first model constructed using lm with no variables
 first_model <-
-  lm(Assault_2020 ~FEMALE_Covid_CASES,
+  lm(Assault_2020 ~1,
      data = final_dataset)
-second_model <-
-  lm(Assault_2020 ~ MALE_Covid_CASES,
-     data = final_dataset)
-third_model <-
-  lm(Assault_2020 ~FEMALE_Covid_CASES+MALE_Covid_CASES,
-     data = final_dataset)
-fourth_model<-
-  lm(Assault_2020 ~FEMALE_Covid_CASES+MALE_Covid_CASES+Sporadic,
-     data = final_dataset)
-fifth_model <-
-  lm(Assault_2020 ~ FEMALE_Covid_CASES+MALE_Covid_CASES+Outbreak_Associated,
-     data = final_dataset)
-sixth_model <-
-  lm(Assault_2020 ~ FEMALE_Covid_CASES+MALE_Covid_CASES+Sporadic+Outbreak_Associated,data = final_dataset)
-seventh_model<-
-   lm(Assault_2020 ~ Sporadic+Outbreak_Associated,data = final_dataset)
-eight_model<-lm(Assault_2020 ~ 1,data = final_dataset)
-ninth_model<-lm(Assault_2020~Total_CASES,data=final_dataset)
-modelsummary(list(first_model,second_model,third_model,fourth_model,fifth_model,sixth_model, seventh_model,eight_model,ninth_model))
+modelsummary(first_model)
 ```
 
-```{r model, echo=FALSE, message=FALSE, include=FALSE}
-final_model<-lm(formula=Assault_2020~FEMALE_Covid_CASES+MALE_Covid_CASES,data=final_dataset)
+The AIC obtained from this model is 1757.7, so we will add a variable to see if we can reduce it.
+
+```{r, echo=TRUE}
+second_model<-
+  lm(Assault_2020 ~FEMALE_Covid_CASES,data = final_dataset)
+modelsummary(second_model)
+```
+
+The AIC obtained from this model is 1745.0 which is less so we can keep the female variable.
+
+```{r, echo=TRUE}
+third_model<-
+  lm(Assault_2020 ~FEMALE_Covid_CASES+MALE_Covid_CASES,data = final_dataset)
+modelsummary(third_model)
+```
+
+AIC obtained is 1728.4 so we can keep going.
+
+```{r, echo=TRUE}
+fourth_model<-
+  lm(Assault_2020 ~FEMALE_Covid_CASES+MALE_Covid_CASES+transgender_covid_cases,data = final_dataset)
+modelsummary(fourth_model)
+```
+
+The AIC obtained is 1729.5 so we don't include transgender covid cases
+
+```{r, echo=TRUE}
+fifth_model<-
+  lm(Assault_2020 ~FEMALE_Covid_CASES+MALE_Covid_CASES+`20 to 29`,data = final_dataset)
+modelsummary(fifth_model)
+```
+
+We keep the 20 to 29 variable since the AIC is lower.
+
+```{r, echo=TRUE}
+sixth_model<-
+  lm(Assault_2020 ~FEMALE_Covid_CASES+MALE_Covid_CASES+`20 to 29`+`30 to 39`,data = final_dataset)
+modelsummary(sixth_model)
+```
+
+So we move on with this model since the AIC is 1697.2
+
+```{r, echo=TRUE}
+seventh_model<-
+  lm(Assault_2020 ~FEMALE_Covid_CASES+MALE_Covid_CASES+`20 to 29`+`30 to 39`+`40 to 49`,data = final_dataset)
+modelsummary(seventh_model)
+```
+
+We proceed on with this model and will try to remove the 20 to 29 to see if we get a lower AIC.
+
+```{r, echo=TRUE}
+eight_model<-
+  lm(Assault_2020 ~FEMALE_Covid_CASES+MALE_Covid_CASES+`30 to 39`+`40 to 49`,data = final_dataset)
+modelsummary(eight_model)
+```
+
+The AIC is not reduced so we keep 20 to 29.
+
+```{r, echo=TRUE}
+seventh_model<-
+  lm(Assault_2020 ~FEMALE_Covid_CASES+MALE_Covid_CASES+`20 to 29`+`40 to 49`,data = final_dataset)
+modelsummary(seventh_model)
+```
+
+This time the AIC is higher so we do not go with this model.
+
+So the final model we obtain is the seventh model
+
+Assault_2020 ~FEMALE_Covid_CASES+MALE_Covid_CASES+`20 to 29`+ 30 to 39`+`40 to 49`
+
+## Model
+
+These are the estimates obtained for the final model using modelsummary.
+
+```{r model, echo=TRUE}
+final_model<-
+  lm(Assault_2020 ~FEMALE_Covid_CASES+MALE_Covid_CASES+`20 to 29`+`30 to 39`+`40 to 49`,
+     data = final_dataset)
 modelsummary(final_model)
@@
 
-# Additional details
+# Appendix B
+
+## Covid Cases
+
+This is a preview of the covid_cases dataset obtained from opendatatoronto [@citeopendatatoronto]
+
+```{r dataset1, echo=FALSE}
+covid_cases<-
+  read_csv(here::here("inputs/data/covid_cases.csv"))
+covid_cases
+```
+
+## Neighbourhood crime rates
+
+This is a preview of the neighbourhood crimes dataset obtained from opendatatoronto [@citeopendatatoronto]
+
+```{r dataset2, echo=FALSE}
+neighbourhood_crime_rates<-
+  read_csv(here::here("inputs/data/neighbourhood_crime_rates.csv"))
+neighbourhood_crime_rates
+```
+
+## Appendix C
+
+The datasheet below is an enhancement for the paper and we will be constructing the datasheet for the covid cases database
+
+**Motivation**
+
+1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
+    - The dataset was created to allow people to track the number of reported covid cases in Toronto and contains demographic, geographic and severity information for all cases of covid that were reported whether or not they were probable or confirmed.
+2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
+    - The dataset was created by the Toronto Public Health organization on behalf of the city. 
+3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
+    - Toronto Public Health organization is a public organization so the creation of the dataset was funded by the city.
+4. *Any other comments?*
+    - NA
+
+**Composition**
+
+1.
+*What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)?
+Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)?
+Please provide a description.*
+	- The instances that comprise the dataset represent the neighbourhoods in Toronto. This includes downtown Toronto, Scarborough neighbourhoods, etc. 
+2. *How many instances are there in total (of each type, if appropriate)?*
+	- There are 140 instances since there are 140 neighbourhoods. 
+3.
+*Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set?
+If the dataset is a sample, then what is the larger set?
+Is the sample representative of the larger set (for example, geographic coverage)?
+If so, please describe how this representativeness was validated/verified.
+If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
+	- The dataset contains all possible instances since it includes all 140 neighbourhoods in TOronto.
+4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
+	- Each instance consists of the details about a reported covid case, whether the case was probably or confirmed, whether the client gender was male, female, trans etc, whether it was from an outbreak or sporadic and whether or not the case was resolved.
+5. *Is there a label or target associated with each instance? If so, please provide a description.*
+	- There is no target associated with any of the instances. 
+6.
+*Is any information missing from individual instances?
+If so, please provide a description, explaining why this information is missing (for example, because it was unavailable).
+This does not include intentionally removed information, but might include, for example, redacted text.*
+	- There are some NA cases where information in the data features is missing, possibly because they were unable to collect that information when it was reported.
+7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
+	- There are no relationships made explicit between instances.
+8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
+	- There are no recommended data splits.
+9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
+	- There are no errors, source of noise or redundancies in the dataset.
+10.
+*Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)?
+If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer?
+Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
+	- TThe dataset updates on a weekly basis since the pandemic is still ongoing. 
+11.
+*Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)?
+If so, please provide a description.*
+	- It does not contain confidential data
+12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
+	- It does not contain such data.
+13.
+*Does the dataset identify any sub-populations (for example, by age, gender)?
+If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
+	- The dataset identifies the gender of the client and the age group of the reported individual with covid. 
+14.
+*Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset?
+If so, please describe how.*
+	- It is not possible
+15.
+*Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)?
+If so, please provide a description.*
+	- It does not contain such sensitive data. 
+16. *Any other comments?*
+	- NA
+
+**Collection process**
+
+1.
+*How was the data associated with each instance acquired?
+Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)?
+If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified?
+If so, please describe how.*
+	- The data was extracted from the provincial Case & Contact Management System (CCM). 
+2.
+*What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)?
+How were these mechanisms or procedures validated?*
+	- The data was collected by recording the features for the person who was reported to have covid in each instance. 
+3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
+	- NA
+4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
+	- The provincial Case & Contact Management System was involved and their compensation is not recorded
+5.
+*Over what timeframe was the data collected?
+Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)?
+If not, please describe the timeframe in which the data associated with the instances was created.*
+	- The data collects all covid cases from January 2020 to the present date.
+6.
+*Were any ethical review processes conducted (for example, by an institutional review board)?
+If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
+	- No ethical review processes appear to have been conducted
+7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
+	- I obtained it using opendatatoronto, a public website. 
+8.
+*Were the individuals in question notified about the data collection?
+If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
+	- No
+9.
+*Did the individuals in question consent to the collection and use of their data?
+If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
+	- Since it was a public website, I did not require their consent.
+10.
+*If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses?
+If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
+	- NA
+11.
+*Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted?
+If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
+	- NA
+12. *Any other comments?*
+	- NA
+
+**Preprocessing/cleaning/labeling**
+
+1.
+*Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)?
+If so, please provide a description.
+If not, you may skip the remaining questions in this section.*
+	- NA
+2.
+*Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)?
+If so, please provide a link or other access point to the "raw" data.*
+	- NA
+3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
+	- NA
+4. *Any other comments?*
+	- NA
+
+**Uses**
+
+1. *Has the dataset been used for any tasks already? If so, please provide a description.*
+	- The dataset has been used for no other tasks yet.
+2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
+	- https://github.com/Varun1005473462/final_folder-main.git
+3. *What (other) tasks could the dataset be used for?*
+	- It could be used for modelling the spread of covid in Toronto neighbourhoods since 2020.
+4.
+*Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses?
+For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)?
+If so, please provide a description.
+Is there anything a dataset consumer could do to mitigate these risks or harms?*
+	- TBD
+5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
+	- TBD
+6. *Any other comments?*
+	- TBD
+
+**Distribution**
+
+1.
+*Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created?
+If so, please provide a description.*
+	- The dataset will not be distributed to third parties outside of the entity on behalf of which the dataset was created
+2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
+	- NA
+3. *When will the dataset be distributed?*
+	- NA
+4.
+*Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)?
+If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
+	- No
+5.
+*Have any third parties imposed IP-based or other restrictions on the data associated with the instances?
+If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
+	- No
+6.
+*Do any export controls or other regulatory restrictions apply to the dataset or to individual instances?
+If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
+	- No
+7. *Any other comments?*
+	- NA
+
+**Maintenance**
+
+1. *Who will be supporting/hosting/maintaining the dataset?*
+	- Varun Vijay
+2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
+	- through mail at varun.vijay@mail.utoronto.ca
+3. *Is there an erratum? If so, please provide a link or other access point.*
+	- No
+4.
+*Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)?
+If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
+	- The dataset is updated weekly by the city to account for new covid cases. Any updates can be communicated through github. 
+5.
+*If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)?
+If so, please describe these limits and explain how they will be enforced.*
+	- No
+6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
+	- No. Its obsolescence is communicated via commit history on Github
+7.
+*If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so?
+If so, please provide a description.
+Will these contributions be validated/verified?
+If so, please describe how.
+If not, why not?
+Is there a process for communicating/distributing these contributions to dataset consumers?
+If so, please provide a description.*
+	- f others want to extend/augment/build on/contribute to the dataset, they can do s using the github link above.
+These contributions will not be validated/ verified since I am not responsible for such data extensions
+8. *Any other comments?*
+	- N/A
 
@@
 # References
+
+<div id = "refs"></div>
+
+[@thereferencecanbewhatever]
+
+[@ggplot2]
+
+[@dplyr]
+
+[@tidyr]
+
+[@janitor]
+
+[@reshape2]
+
+[@here]
+
+[@tinytex]
+
+[@readr]
+
+[@metaopendata1]
+
+[@metaopendata2]

### 2022-04-27 outputs/paper/paper.Rmd
Completed paper, finished discussion and proofread. completed datasheet and knitted pdf

@@
 x_2 = number of male covid cases in a neighbourhood,
-x_3 = number of covid cases for people between ages 20 and 29
+x_3 = number of covid cases for people between ages 20 and 29,
 x_4 = number of covid cases for people between ages 30 and 39,

### 2022-04-27 outputs/paper/paper.Rmd
completed proofreading of the rmd file and finished final changes

@@
 The dataset includes all cases related to outbreaks and sporadic cases, those that occur in the community.
-The data was extracted from the provincial Case & Contact Management System and the variables in the dataset include the neighbourhood the case was reported in, the date it was reported, the client gender, Age group and the classification of the case, whether it was probable or improbable.
+The data was collected by opendatatoronto from the provincial Case & Contact Management System and the variables in the dataset include the neighbourhood the case was reported in, the date it was reported, the client gender, Age group and the classification of the case, whether it was probable or improbable.
 The full dataset \@ref(table: dataset1) can be found in Appendix B.
@@
 We then focused on the variables Client Gender and the age group of the infected.
-These variables are the ones that pertain the most to the total number of covid cases in a neighbourhood and so are the ones we will be focusing on for this dataset as they also more important than the others.
-New variables using the dataset were created such as number of female covid cases, number of male covid cases, number of transgender covid cases, number of other genders and unknown genders covid cases and variables related to the various age groups, less than 19, between 20 to 29, 30 to 39, etc.
+These variables are the ones that pertain the most to the total number of covid cases in a neighbourhood and so are the ones we will be focusing on for this dataset.
+New variables were created such as number of female covid cases, number of male covid cases, number of transgender covid cases, number of other genders and unknown genders covid cases and variables related to the various age groups, less than 19, between 20 to 29, 30 to 39, etc.
 We also created a new variable where we counted the total number of Covid cases in a neighbourhood.
@@
 All NA cases were also deleted in order to improve the analysis.
+All of these variables were placed together in a dataset for analysis.
+This dataset is the covid dataset we will be using.
 
 For the crime dataset, we only focused on the neighbourhoods and the number of assaults for the year 2020 since these pertain directly to my paper.
-We then joined the two datasets together and obtained the final_dataset.
+We then joined the crime and covid datasets together and obtained the final_dataset.
 All 140 neighbourhoods are included in this dataset and it is the dataset we will be using for the rest of this paper.
@@
 From the above table \@ref(table: summary 2) we can see that the variables with the strongest correlation with assaults in 2020 are number of female covid cases, number of male covid cases, number of covid cases for age groups 20 to 29, 30 to 39 and 40 to 49.
-It makes for these variables to relate to number of assaults in a neighbourhood since they are the most important ones since age and gender are incredibly important in such situations when an assault is occurring.
+It makes sense for these variables to relate to number of assaults in a neighbourhood since they are the most important ones since age and gender are incredibly important when building a model for crime and assaults.
 We also see that age groups and the number of male and female covid cases are strongly related.
@@
 Fig \@ref(fig: figure 1) shows us the number of assaults in 2020 per neighbourhood.
-There does seem to be some slight variation in the number of assaults per neighbourhood with some high outliers which match the results shown in the summary table.
+There does seem to be some variation in the number of assaults per neighbourhood with some high outliers.
+The plot obtained matches the results shown in the summary table.
 
@@
 
-Figure 4,5,6,7,8 and 9 all show that there seems to be a positive relation between the number of covid cases in a neighbourhood and the number of assaults in that neighbourhood except for a few outliers.
+Figure 4 \ref(fig:figure 4),5 \ref(fig:figure 5),6 \ref(fig:figure 6),7 \ref(fig:figure 7),8 \ref(fig:figure 8) and 9 \ref(fig:figure 9) all show that there seems to be a positive relation between the number of covid cases in a neighbourhood and the number of assaults in that neighbourhood except for a few outliers.
 When the number of covid cases in a neighbourhood are low so are the number of assaults and when the number of covid cases increase so do the number of assaults increase with it.
@@
 
-We can analyse the goodness of fit for the model from the above 4 plots \ref(fig:figure 10).
+We can analyse the goodness of fit for the model from the above 4 plots in figure 10 \ref(fig:figure 10).
 From the residuals vs fitted plot, we see that linearity for the model does hold and that the independent variables do enter in a linear way.
@@
 
-We proceed on with this model and will try to remove the 20 to 29 to see if we get a lower AIC.
+We proceed with this model and will try to remove the 20 to 29 to see if we get a lower AIC.
 
@@
 
-The AIC is not reduced so we keep 20 to 29.
+The AIC is not reduced so we keep 20 to 29 and we try and remove another variable.
 
